---
title: "NOAA Temperature Data"
subtitle: "Data for Eastern Shore Region"
author: "Tolu Odukoya"
date: "7/6/2021"
output: 
  html_document: 
    toc: true
    toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
# 0. Load libraries and data
library(tidyverse)
library(stargazer) # for summary table
library(janitor) # for tabyl
library(tigris) # for shapefiles
library(sf) # for spatial joins
library(leaflet) # for map
library(viridis)
library(googlesheets4)
library(RColorBrewer)
library(ggplot2)
library(gganimate)
library(gapminder)
library(transformr)
library(gifski)
library(ggrepel)
library(hrbrthemes)
library(devtools)
library(climatestripes)
library(lubridate)
noaa <- read_csv("noaa_eastern_county.csv")
easternfips <- c("001", "131")
meta <- read_sheet("https://docs.google.com/spreadsheets/d/1nqm3DuVXD1ObbVe_deacvT7uSLdBXfQJo3mkbqDwrVo/edit#gid=1573436636", sheet = "noaa")
```

```{r load and clean new data shp, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

eastern_sf <- read_sf("/Users/msbugatti/Documents/Intership 2021/NOAA Temp Work/counties")
eastern_sf <- eastern_sf %>% filter(FIPS %in% c(51001, 51131))
eastern_sf <- eastern_sf %>%
                      dplyr::select(NAME, STATE_FIPS, CNTY_FIPS, FIPS, geometry, AREA)
eastern_sf1 <- left_join(eastern_sf, noaa, by ="CNTY_FIPS")
```


```{r transpose the data, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

eastern_sf1 <- st_as_sf(eastern_sf1)
eastern_sf1 <- cbind(eastern_sf1, st_coordinates(st_centroid(eastern_sf1)))
```

```{r remove na, , include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

eastern_sf1 <- eastern_sf1 %>%
                      dplyr::select(NAME, Year, CNTY_FIPS, X, Y,
                                geometry, everything())
eastern_sf1$Year = as.numeric(eastern_sf1$Year)
eastern_sf1 <-  eastern_sf1 %>% filter_at(vars(Janmin:Avg_Temppcp),all_vars(!is.na(.)))
```

## Data Source

Source: NOAA's Climate Divisional Database (nClimDiv), June 2021 release

* Download URL: [https://www.ncei.noaa.gov/pub/data/cirs/climdiv/counties.zip](https://www.ncei.noaa.gov/pub/data/cirs/climdiv/counties.zip)
* Download URL: [https://www.ncei.noaa.gov/pub/data/cirs/climdiv/climdiv-tmaxcy-v1.0.0-20210604](https://www.ncei.noaa.gov/pub/data/cirs/climdiv/climdiv-tmaxcy-v1.0.0-20210604)

### The Data Used Here

The Climate Divisional Dataset from the National Center For Environmental Information began as the only long-term temporally and spatially complete dataset from which to generate historical climate analyses (1895-2013) for the contiguous United States (CONUS). It was originally developed for climate division, statewide, regional, national, and population-weighted monitoring of drought, temperature, precipitation, and heating/cooling degree day values. 

There are 344 climate divisions in the CONUS. For each climate division, monthly station temperature and precipitation values are computed from the daily observations. The divisional values are weighted by area to compute statewide values and the statewide values are weighted by area to compute regional values. (Karl and Koss, 1984).

The data here is the county data from all the states in the contigous United States. While NOAA includes information on a variety of climate measures, we are focusing on the maximum, minimum temperature, and precipitation of the counties of interest. 

To Learn More, See: 

* [ National Center For Environmental Information and Data ](https://www.ncdc.noaa.gov/monitoring-references/maps/us-climate-divisions.php#grdd)


## Variable Descriptions

```{r}
glimpse(eastern_sf1)
```

Observations are county level estimates of...

* Monthly maximum temperature for each county for years 1895-2021
    * Yearly average maximum temperature for each county for those years. 
    

* Monthly minimum temperature for each county for years 1895-2021
    * Yearly average minimum temperature for each county for those years. 


* Monthly precipitation for each county for years 1895-2021
    * Yearly average precipitation for each county for those years.


## Summaries 

5-number summaries of (non-missing) numeric variables (remove non-numeric observations)

```{r}
noaa %>% select(-c(CNTY_FIPS, Fipsyear, Year)) %>% 
  select(where(~is.numeric(.x) && !is.na(.x))) %>% 
  as.data.frame() %>% 
  stargazer(., type = "text", title = "Summary Statistics", digits = 0,
            summary.stat = c("mean", "sd", "min", "median", "max"))
```


## Maps

### Maximum Temperature 

#### Maximum Temperature in July across Counties for all Years

```{r}
eastern_sf1a <- 
  ggplot(eastern_sf1) +
  geom_sf(aes(fill = Julmax), color = "black", alpha = .9, na.rm = TRUE) +
   geom_text_repel(data = eastern_sf1, aes(X, Y, label = NAME), size = 4, nudge_x = 1, nudge_y = 0, fontface = "bold", hjust = 0.9) +
  scale_fill_fermenter(palette = "YlOrRd", direction = 1,   type = "seq", n.breaks = 7) +
     theme_void() +
  guides(fill =
           guide_colourbar(title.position="top", title.hjust = 0.5,
                           barwidth = 1)
  )  + 
  labs(fill = "Temperature ", title = 'Year: {frame_time}',
       caption = "Maximum Temperature in July for Eastern Shore Counties") + 
 transition_time(as.integer(Year)) +
ease_aes('linear') 
```

```{r}
animate(eastern_sf1a, fps = 1, detail = 1, nframes = 127)
```

```{r, results = "asis"}
meta %>%
  filter(varname == "Julmax") %>% 
  select(about) %>% 
  as.list()
```

#### Warming Stripes Showing Maximum Temperature in July for Each County for all Years 
```{r warming1, echo = TRUE, warning = FALSE, message = FALSE}

#select only the july temperature and year column
eastern_sf1_jl <- select(eastern_sf1, NAME, Year, Julmax)

#rename the temperature column
eastern_sf1_jl <- rename(eastern_sf1_jl, ta = Julmax)

```

```{r warming2, echo = TRUE, warning = FALSE, message = FALSE}

#create a date column because stripes only works with format = date

eastern_sf1_jl <- mutate(eastern_sf1_jl, date = str_c(Year, "01-01", sep = "-") %>% ymd())

```

```{r  warming3, echo = TRUE, warning = FALSE, message = FALSE}

#Filter out each County
eastern_sf1_jla <- filter(eastern_sf1_jl, NAME == "Accomack")
eastern_sf1_jln <- filter(eastern_sf1_jl, NAME == "Northampton")
```

```{r  warming4,  echo = TRUE, warning = FALSE, message = FALSE}

# Create a theme for the stripes image

theme_strip <- theme_minimal()+
                 theme(axis.text.y = element_blank(),
                       axis.line.y = element_blank(),
                       axis.title = element_blank(),
                       panel.grid.major = element_blank(),
                       legend.title = element_blank(),
                       axis.text.x = element_text(vjust = 3),
                       panel.grid.minor = element_blank(),
                        plot.title = element_text(size = 14, face = "bold")
                       )


col_strip <- brewer.pal(11, "RdBu")
```

##### Accomack County Warming Stripes: July Maximum Temperature
```{r stripes1,fig.width=10, fig.height=3, message=FALSE, warning=FALSE}
 ggplot(eastern_sf1_jla,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Accomack County July Maximum Temperature 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```

##### Northampton County Warming Stripes: July Maximum Temperature
```{r stripes2,fig.width=10, fig.height=3, message=FALSE, warning=FALSE}
 ggplot(eastern_sf1_jln,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Northampton County July Maximum Temperature 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```


### Average Yearly Maximum Temperature across Counties

```{r}
eastern_sf1b <- 
  ggplot(eastern_sf1) +
  geom_sf(aes(fill = Avg_Tempmax), color = "black", alpha = .9, na.rm = TRUE) +
   geom_text_repel(data = eastern_sf1, aes(X, Y, label = NAME), size = 4, nudge_x = 1, nudge_y = 0, fontface = "bold", hjust = 0.9) +
  scale_fill_fermenter(palette = "BuPu", direction = 1,   type = "seq", n.breaks = 6) +
     theme_void() +
  guides(fill =
           guide_colourbar(title.position="top", title.hjust = 0.5,
                           barwidth = 1)
  )  + 
  labs(fill = "Temperature ", title = 'Year: {frame_time}',
       caption = "Average Yearly Maximum Temperature For Eastern Shore Counties") + 
 transition_time(as.integer(Year)) +
ease_aes('linear') 
```

```{r}
animate(eastern_sf1b, fps = 1, detail = 1, nframes = 127)
```


```{r, results = "asis"}
meta %>%
  filter(varname == "Avg_Tempmax") %>% 
  select(about) %>% 
  as.list()
```

#### Warming Stripes Showing Average Yearly Maximum Temperature for Each County
```{r  warming6, echo = TRUE, warning = FALSE, message = FALSE}

#select only the annual temperature and year column
eastern_sf1_yr <- select(eastern_sf1, NAME, Year, Avg_Tempmax)

#rename the temperature column
eastern_sf1_yr <- rename(eastern_sf1_yr, ta = Avg_Tempmax)

```

```{r  warming7, echo = TRUE, warning = FALSE, message = FALSE}

#create a date column because stripes only works with format = date

eastern_sf1_yr <- mutate(eastern_sf1_yr, date = str_c(Year, "01-01", sep = "-") %>% ymd())

```

```{r warming8, echo = TRUE, warning = FALSE, message = FALSE}
#Filter out each County. No need to do theme again it's already set 

eastern_sf1_yra <- filter(eastern_sf1_yr, NAME == "Accomack")
eastern_sf1_yrn <- filter(eastern_sf1_yr, NAME == "Northampton")
```

##### Accomack County Warming Stripes: Average Yearly Maximum Temperature
```{r stripes1a,fig.width=10, fig.height=3, message=FALSE, warning=FALSE}
 ggplot(eastern_sf1_yra,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Accomack County Average Yearly Maximum Temperature 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```

##### Northampton County Warming Stripes: Average Yearly Maximum Temperature
```{r stripes2a,fig.width=10, fig.height=3, message=FALSE, warning=FALSE}
 ggplot(eastern_sf1_yrn,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Northampton County Average Yearly Maximum Temperature 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```



### Minimum Temperature

#### Minimum Temperature in July across Counties for all Years 

```{r}
eastern_sf2a <- 
  ggplot(eastern_sf1) +
  geom_sf(aes(fill = Julmin), color = "black", alpha = .9, na.rm = TRUE) +
   geom_text_repel(data = eastern_sf1, aes(X, Y, label = NAME), size = 4, nudge_x = 1, nudge_y = 0, fontface = "bold", hjust = 0.9) +
  scale_fill_fermenter(palette = "YlGnBu", direction = 1,   type = "seq", n.breaks = 7) +
     theme_void() +
  guides(fill =
           guide_colourbar(title.position="top", title.hjust = 0.5,
                           barwidth = 1)
  )  + 
  labs(fill = "Temperature ", title = 'Year: {frame_time}',
       caption = "Minimum Temperature in July for Eastern Shore Counties") + 
 transition_time(as.integer(Year)) +
ease_aes('linear') 
```

```{r}
animate(eastern_sf2a, fps = 1, detail = 1, nframes = 127)
```

```{r, results = "asis"}
meta %>%
  filter(varname == "Julmin") %>% 
  select(about) %>% 
  as.list()
```

#### Warming Stripes Showing Minimum Temperature in July for Each County for all Years 
```{r warming1b, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

#select only the july temperature and year column
eastern_sf2_jl <- select(eastern_sf1, NAME, Year, Julmin)

#rename the temperature column
eastern_sf2_jl <- rename(eastern_sf2_jl, ta = Julmin)

```

```{r warming2b, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

eastern_sf2_jl <- mutate(eastern_sf2_jl, date = str_c(Year, "01-01", sep = "-") %>% ymd())

```

```{r  warming3b, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

eastern_sf2_jla <- filter(eastern_sf2_jl, NAME == "Accomack")
eastern_sf2_jln <- filter(eastern_sf2_jl, NAME == "Northampton")
```


##### Accomack County Warming Stripes: July Minimum Temperature
```{r stripes1b,fig.width=10, fig.height=3, message=FALSE, warning=FALSE}
 ggplot(eastern_sf2_jla,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Accomack County July Minimum Temperature 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```

##### Northampton County Warming Stripes: July Minimum Temperature
```{r stripes2b,fig.width=10, fig.height=3, message=FALSE, warning=FALSE}
 ggplot(eastern_sf2_jln,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Northampton County July Minimum Temperature 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```


### Average Yearly Minimum Temperature across Counties

```{r}
eastern_sf2b <- 
  ggplot(eastern_sf1) +
  geom_sf(aes(fill = Avg_Tempmin), color = "black", alpha = .9, na.rm = TRUE) +
   geom_text_repel(data = eastern_sf1, aes(X, Y, label = NAME), size = 4, nudge_x = 1, nudge_y = 0, fontface = "bold", hjust = 0.9) +
  scale_fill_fermenter(palette = "PuRd", direction = 1,   type = "seq", n.breaks = 6) +
     theme_void() +
  guides(fill =
           guide_colourbar(title.position="top", title.hjust = 0.5,
                           barwidth = 1)
  )  + 
  labs(fill = "Temperature ", title = 'Year: {frame_time}',
       caption = "Average Yearly Minimum Temperature For Eastern Shore Counties") + 
 transition_time(as.integer(Year)) +
ease_aes('linear') 
```

```{r}
animate(eastern_sf2b, fps = 1, detail = 1, nframes = 127)
```


```{r, results = "asis"}
meta %>%
  filter(varname == "Avg_Tempmin") %>% 
  select(about) %>% 
  as.list()
```

#### Warming Stripes Showing Average Yearly Minimum Temperature for Each County
```{r  warming6a, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

#select only the annual temperature and year column
eastern_sf2_yr <- select(eastern_sf1, NAME, Year, Avg_Tempmin)

#rename the temperature column
eastern_sf2_yr <- rename(eastern_sf2_yr, ta = Avg_Tempmin)

```

```{r  warming7a, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

eastern_sf2_yr <- mutate(eastern_sf2_yr, date = str_c(Year, "01-01", sep = "-") %>% ymd())

```

```{r warming8a, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

eastern_sf2_yra <- filter(eastern_sf2_yr, NAME == "Accomack")
eastern_sf2_yrn <- filter(eastern_sf2_yr, NAME == "Northampton")
```

##### Accomack County Warming Stripes: Average Yearly Minimum Temperature
```{r stripes1c,fig.width=10, fig.height=3, message=FALSE, warning=FALSE}
 ggplot(eastern_sf2_yra,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Accomack County Average Yearly Minimum Temperature 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```

##### Northampton County Warming Stripes: Average Yearly Minimum Temperature
```{r stripes2c,fig.width=10, fig.height=3, message=FALSE, warning=FALSE}
 ggplot(eastern_sf2_yrn,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Northampton County Average Yearly Minimum Temperature 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```



### Precipitation

#### Precipitation in July across Counties for all Years 

```{r}
eastern_sf3a <- 
  ggplot(eastern_sf1) +
  geom_sf(aes(fill = Julpcp), color = "black", alpha = .9, na.rm = TRUE) +
   geom_text_repel(data = eastern_sf1, aes(X, Y, label = NAME), size = 4, nudge_x = 1, nudge_y = 0, fontface = "bold", hjust = 0.9) +
  scale_fill_fermenter(palette = "PuBuGn", direction = 1,   type = "seq", n.breaks = 7) +
     theme_void() +
  guides(fill =
           guide_colourbar(title.position="top", title.hjust = 0.5,
                           barwidth = 1)
  )  + 
  labs(fill = "Precipitation", title = 'Year: {frame_time}',
       caption = "Precipitation in July for Eastern Shore Counties") + 
 transition_time(as.integer(Year)) +
ease_aes('linear') 
```

```{r}
animate(eastern_sf3a, fps = 1, detail = 1, nframes = 127)
```

```{r, results = "asis"}
meta %>%
  filter(varname == "Julpcp") %>% 
  select(about) %>% 
  as.list()
```

#### Warming Stripes Showing Precipitation in July for Each County for all Years 
```{r warming1d, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

#select only the july temperature and year column
eastern_sf3_jl <- select(eastern_sf1, NAME, Year, Julpcp)

#rename the temperature column
eastern_sf3_jl <- rename(eastern_sf3_jl, ta = Julpcp)

```

```{r warming2d, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

eastern_sf3_jl <- mutate(eastern_sf3_jl, date = str_c(Year, "01-01", sep = "-") %>% ymd())

```

```{r  warming3d, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

eastern_sf3_jla <- filter(eastern_sf3_jl, NAME == "Accomack")
eastern_sf3_jln <- filter(eastern_sf3_jl, NAME == "Northampton")
```


##### Accomack County Warming Stripes: July Precipitation
```{r stripes1d,fig.width=10, fig.height=3, message=FALSE, warning=FALSE}
 ggplot(eastern_sf3_jla,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Accomack County July Precipitation 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```

##### Northampton County Warming Stripes: July Precipitation
```{r stripes2f,fig.width=10, fig.height=3, message=FALSE, warning=FALSE}
 ggplot(eastern_sf3_jln,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Northampton County July Precipitation 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```


### Average Yearly Precipitation across Counties

```{r}
eastern_sf3b <- 
  ggplot(eastern_sf1) +
  geom_sf(aes(fill = Avg_Temppcp), color = "black", alpha = .9, na.rm = TRUE) +
   geom_text_repel(data = eastern_sf1, aes(X, Y, label = NAME), size = 4, nudge_x = 1, nudge_y = 0, fontface = "bold", hjust = 0.9) +
  scale_fill_fermenter(palette = "RdPu", direction = 1,   type = "seq", n.breaks = 6) +
     theme_void() +
  guides(fill =
           guide_colourbar(title.position="top", title.hjust = 0.5,
                           barwidth = 1)
  )  + 
  labs(fill = "Precipitation", title = 'Year: {frame_time}',
       caption = "Average Yearly Precipitation For Eastern Shore Counties") + 
 transition_time(as.integer(Year)) +
ease_aes('linear') 
```

```{r}
animate(eastern_sf3b, fps = 1, detail = 1, nframes = 127)
```


```{r, results = "asis"}
meta %>%
  filter(varname == "Avg_Temppcp") %>% 
  select(about) %>% 
  as.list()
```

#### Warming Stripes Showing Average Yearly Precipitation for Each County
```{r  warming6d, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

#select only the annual temperature and year column
eastern_sf3_yr <- select(eastern_sf1, NAME, Year, Avg_Temppcp)

#rename the temperature column
eastern_sf3_yr <- rename(eastern_sf3_yr, ta = Avg_Temppcp)

```

```{r  warming7d, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

eastern_sf3_yr <- mutate(eastern_sf3_yr, date = str_c(Year, "01-01", sep = "-") %>% ymd())

```

```{r warming8d, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

eastern_sf3_yra <- filter(eastern_sf3_yr, NAME == "Accomack")
eastern_sf3_yrn <- filter(eastern_sf3_yr, NAME == "Northampton")
```

##### Accomack County Warming Stripes: Average Yearly Precipitation
```{r stripes1f,fig.width=10, fig.height=3, message=FALSE, warning=FALSE}
 ggplot(eastern_sf3_yra,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Accomack County Average Yearly Precipitation 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```

##### Northampton County Warming Stripes: Average Yearly Precipitation
```{r stripes2g,fig.width=10, fig.height=3, message=FALSE, warning=FALSE}
 ggplot(eastern_sf3_yrn,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Northampton County Average Yearly Precipitation 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```

