---
title: "Air Quality Exploration File"
subtitle: "Data for Charlottesville Region"
author: "Marisa Lemma, Michele Claibourn"
date: "11/12/2021"
output: 
  html_document: 
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# Load libraries and data
library(tidyverse)
library(stargazer)
library(leaflet)
library(sf)
library(googlesheets4)
library(plotly)

airquality <- read_csv("../data/airquality_cville_tract.csv")
airquality <- airquality %>% 
  mutate(gid = as.character(gid))

meta <- read_sheet("https://docs.google.com/spreadsheets/d/1nqm3DuVXD1ObbVe_deacvT7uSLdBXfQJo3mkbqDwrVo/edit#gid=627666074", sheet="airquality", gs4_deauth())

```

## Data Source

Source: Replication Data for "Disparities in PM~2.5~ air pollution in the United States"

* Download URL: https://dataverse.lib.virginia.edu/dataset.xhtml?persistentId=doi:10.18130/V3/QLBNQJ
  + Article and Findings: https://science-sciencemag-org.proxy01.its.virginia.edu/content/369/6503/575
* Additional Information: https://science-sciencemag-org.proxy01.its.virginia.edu/content/sci/suppl/2020/07/29/369.6503.575.DC1/aaz9353_Colmer_SM.pdf
* Data include estimates from 1981 through 2016. No updates are planned.


## About the Data

These data show air pollution, via the concentrations of fine particulate matter that is less than 2.5 micrometers in diameter (PM~2.5~), at each census tract. PM~2.5~ concentrations are measured by the number of micrograms per cubic meter. High concentrations of PM~2.5~ indicate higher levels of air pollution. 

The data provides PM~2.5~ concentrations at every year from 1981-2016. 


## Variable Descriptions
```{r}
glimpse(select(airquality, statefp, countyfp, tract, 
               pm2_5_1981, pm2_5_2016, pm_change_1981_2016,
               percentile_1981, percentile_2016,
               pctile_change_1981_2016))
```

Observations are census tract estimates of:

* PM~2.5~ levels in 1981 through 2016 (`pm2_5_1981`-`pm2_5_2016`)
* Percentile rankings in 1981 and 2016 (`percentile_1981` and `percentile_2016`)
  + Percentile rankings were calculated among census tracts within the Commonwealth and were not altered once the data were filtered to just the Charlottesville region.
* Change in PM~2.5~ level between 1981 and 2016 (`pm_change_1981_2016`)
* Change in percentile rank between 1981 and 2016 (`pctile_change_1981_2016`)

## Summaries

Five-number summaries of all variables:
```{r}
airquality %>% select(statefp, countyfp, tract, 
               pm2_5_1981, pm2_5_2016, pm_change_1981_2016,
               percentile_1981, percentile_2016,
               pctile_change_1981_2016) %>% 
  as.data.frame() %>% 
  stargazer(., type = "text", title = "Summary Statistics", digits = 0,
            summary.stat = c("mean", "sd", "min", "median", "max"))
```

## Visual Distributions

### Tract distributions of PM~2.5~ in 1981 and 2016:

```{r, message=F, warning=F}
airquality %>% select(tract, pm2_5_1981, pm2_5_2016) %>% 
  pivot_longer(-tract, names_to = "measure", values_to = "value") %>% 
  ggplot(aes(x = value, fill = measure)) + 
  geom_histogram() + 
  facet_wrap(~measure, scales = "free") +
  xlab("PM2.5") +
  scale_fill_discrete(labels = c("PM2.5 in 1981", "PM2.5 in 2016"))
```

```{r}
meta %>% 
  filter(varname %in% c("pm2_5_1981", "pm2_5_2016")) %>%
  mutate(label = paste0(varname, ": ", about)) %>% 
  select(label) %>% 
  as.list()
```

### Tract-level change from 1981-2016
```{r}
p <- airquality %>% 
  select(-c(percentile_1981, percentile_2016, 
            pm_change_1981_2016, pctile_change_1981_2016)) %>% 
  pivot_longer(cols = starts_with("pm2"), 
               names_to = "year", values_to = "pm2_5",
               names_prefix = "pm2_5_") %>% 
  ggplot(aes(x = year, y = pm2_5, color = countyfp)) +
  geom_line(aes(group = tract)) +
  guides(color = "none")
ggplotly(p) %>% layout(showlegend = FALSE)
```

### Percentile in 1981 vs. percentile in 2016
```{r}
airquality %>% 
  ggplot() +
  geom_point(aes(x=percentile_1981, y=percentile_2016)) +
  xlim(0, 100) +
  ylim(0, 100) +
  geom_abline(intercept = 0, slope = 1, color = "red") +
  xlab("Percentile in 1981") +
  ylab("Percentile in 2016")
```

```{r}
meta %>% 
  filter(varname %in% c("percentile_1981", "percentile_2016")) %>%
  mutate(label = paste0(varname, ": ", about)) %>% 
  select(label) %>% 
  as.list()
```

This scatterplot shows the relationship between a census tract's percentile rank in 1981 and its percentile rank in 2016. The red line shows where the data would be if their percentiles in 1981 and 2016 were the same.


## Spatial Distributions{.tabset}

```{r, include = F}
cville_tracts <- readRDS("../data/cville_tracts.RDS")

cvilleshapes <- left_join(cville_tracts, airquality, by = c('GEOID'= "gid"))
cvilleshapes <- st_transform(cvilleshapes, crs = 4326) # to WGS84, given error
```

### 1981{.tabset}

#### PM~2.5~ Concentration
```{r}
pal <- colorNumeric("Blues", reverse = FALSE, domain = cvilleshapes$pm2_5_1981)

leaflet(cvilleshapes) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = cvilleshapes,
              fillColor = ~pal(pm2_5_1981),
              weight = 1,
              opacity = 1,
              color = "white",
              fillOpacity = 0.6,
              highlight = highlightOptions(weight = 2, fillOpacity = 0.8, bringToFront = T),
              popup = paste0("Tract Number: ", cvilleshapes$GEOID, "<br>",
                             "Concentration: ", cvilleshapes$pm2_5_1981)) %>%
  addLegend("bottomright", pal = pal, values = cvilleshapes$pm2_5_1981,
            title = "PM2.5 Concentration, 1981", opacity = 0.7)
```

```{r}
meta %>% 
  filter(varname=="pm2_5_1981") %>%
  mutate(label = paste0(varname, ": ", about)) %>% 
  select(label) %>% 
  as.list()
```


#### Percentile
```{r}
pal <- colorNumeric("Blues", reverse = FALSE, domain = cvilleshapes$percentile_1981)

leaflet(cvilleshapes) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = cvilleshapes,
              fillColor = ~pal(percentile_1981),
              weight = 1,
              opacity = 1,
              color = "white",
              fillOpacity = 0.6,
              highlight = highlightOptions(weight = 2, fillOpacity = 0.8, bringToFront = T),
              popup = paste0("Tract Number: ", cvilleshapes$GEOID, "<br>",
                             "Percentile: ", round(cvilleshapes$percentile_1981, 2))) %>%
  addLegend("bottomright", pal = pal, values = cvilleshapes$percentile_1981,
            title = "Percentile, 1981", opacity = 0.7)
```

```{r}
meta %>% 
  filter(varname=="percentile_1981") %>%
  mutate(label = paste0(varname, ": ", about)) %>% 
  select(label) %>% 
  as.list()
```

### 2016{.tabset}

#### PM~2.5~ Concentration
```{r}
pal <- colorNumeric("Blues", reverse = FALSE, domain = cvilleshapes$pm2_5_2016)

leaflet(cvilleshapes) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = cvilleshapes,
              fillColor = ~pal(pm2_5_2016),
              weight = 1,
              opacity = 1,
              color = "white",
              fillOpacity = 0.6,
              highlight = highlightOptions(weight = 2, fillOpacity = 0.8, bringToFront = T),
              popup = paste0("Tract Number: ", cvilleshapes$GEOID, "<br>",
                             "Concentration: ", cvilleshapes$pm2_5_2016)) %>%
  addLegend("bottomright", pal = pal, values = cvilleshapes$pm2_5_2016,
            title = "PM2.5 Concentration, 2016", opacity = 0.7)
```

```{r}
meta %>% 
  filter(varname=="pm2_5_2016") %>%
  mutate(label = paste0(varname, ": ", about)) %>% 
  select(label) %>% 
  as.list()
```

#### Percentile
```{r}
pal <- colorNumeric("Blues", reverse = FALSE, domain = cvilleshapes$percentile_2016)

leaflet(cvilleshapes) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = cvilleshapes,
              fillColor = ~pal(percentile_2016),
              weight = 1,
              opacity = 1,
              color = "white",
              fillOpacity = 0.6,
              highlight = highlightOptions(weight = 2, fillOpacity = 0.8, bringToFront = T),
              popup = paste0("Tract Number: ", cvilleshapes$GEOID, "<br>",
                             "Percentile: ", round(cvilleshapes$percentile_2016, 2))) %>%
  addLegend("bottomright", pal = pal, values = cvilleshapes$percentile_2016,
            title = "Percentile, 2016", opacity = 0.7)
```

```{r}
meta %>% 
  filter(varname=="percentile_2016") %>%
  mutate(label = paste0(varname, ": ", about)) %>% 
  select(label) %>% 
  as.list()
```

### Change{.tabset}

#### Change in PM~2.5~, 1981-2016
```{r}
pal <- colorNumeric("Blues", reverse = TRUE, domain = cvilleshapes$pm_change_1981_2016)

leaflet(cvilleshapes) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = cvilleshapes,
              fillColor = ~pal(pm_change_1981_2016),
              weight = 1,
              opacity = 1,
              color = "white",
              fillOpacity = 0.6,
              highlight = highlightOptions(weight = 2, fillOpacity = 0.8, bringToFront = T),
              popup = paste0("Tract Number: ", cvilleshapes$GEOID, "<br>",
                             "PM2.5 Change: ", round(cvilleshapes$pm_change_1981_2016, 2))) %>%
  addLegend("bottomright", pal = pal, values = cvilleshapes$pm_change_1981_2016,
            title = "Change in PM2.5, 1981-2016", opacity = 0.7)
```

```{r}
meta %>% 
  filter(varname=="pm_change_1981_2016") %>%
  mutate(label = paste0(varname, ": ", about)) %>% 
  select(label) %>% 
  as.list()
```

#### Percentile Change, 1981-2016
```{r}
pal <- colorNumeric("Blues", reverse = TRUE, domain = cvilleshapes$pctile_change_1981_2016)

leaflet(cvilleshapes) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = cvilleshapes,
              fillColor = ~pal(pctile_change_1981_2016),
              weight = 1,
              opacity = 1,
              color = "white",
              fillOpacity = 0.6,
              highlight = highlightOptions(weight = 2, fillOpacity = 0.8, bringToFront = T),
              popup = paste0("Tract Number: ", cvilleshapes$GEOID, "<br>",
                             "Percentile Change: ", round(cvilleshapes$pctile_change_1981_2016, 2))) %>%
  addLegend("bottomright", pal = pal, values = cvilleshapes$pctile_change_1981_2016,
            title = "Percentile Change, 1981-2016", opacity = 0.7)
```

```{r}
meta %>% 
  filter(varname=="pctile_change_1981_2016") %>%
  mutate(label = paste0(varname, ": ", about)) %>% 
  select(label) %>% 
  as.list()
```


## Important Notes

The original data uses 2000 census tracts, since that is roughly the midpoint of their 1981-2016 time frame. To integrate this with other 2010 tract-level data, we interpolated the 2000 tract measures to 2010 tracts using areal interpoloation. 
