---
title: "Air Quality Exploration File"
subtitle: "Data for Charlottesville Region"
author: "Marisa Lemma"
date: "7/6/2021"
output: 
  html_document: 
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load libraries and data
library(tidyverse)
library(mosaic)
library(stargazer)
library(viridis)
library(leaflet)
library(rgdal)
library(sf)
library(googlesheets4)

airquality <- read_csv("../data/airquality_cville_tract.csv")

meta <- read_sheet("https://docs.google.com/spreadsheets/d/1nqm3DuVXD1ObbVe_deacvT7uSLdBXfQJo3mkbqDwrVo/edit#gid=627666074", sheet="airquality", gs4_deauth())

```

## Data Source

Source: Replication Data for "Disparities in PM~2.5~ air pollution in the United States"

* Download URL: https://dataverse.lib.virginia.edu/dataset.xhtml?persistentId=doi:10.18130/V3/QLBNQJ
  + Article and Findings: https://science-sciencemag-org.proxy01.its.virginia.edu/content/369/6503/575
* Additional Information: https://science-sciencemag-org.proxy01.its.virginia.edu/content/sci/suppl/2020/07/29/369.6503.575.DC1/aaz9353_Colmer_SM.pdf


## About the Data

This data shows air pollution at each census tract. It specifically focuses on concentrations of PM~2.5~, meaning fine particulate matter that is less than 2.5 micrometers in diameter. PM~2.5~ concentrations are measured by the number of micrograms per cubic meter. It shows PM~2.5~ concentrations at every year from 1981-2016. However, we focus only on 1981 and 2016 and the changes between the two years.


## Variable Descriptions
```{r}
glimpse(airquality)
```

Observations are census tract estimates of:

* PM~2.5~ levels in 1981 and 2016 (`PM2_5_1981` and `PM2_5_2016`)
* Percentile rankings in 1981 and 2016 (`percentile_1981` and `percentile_2016`)
* Change in PM~2.5~ level between 1981 and 2016 (`PM_change`)
* Change in percentile rank between 1981 and 2016 (`pctile_change`)

## Summaries

Five-number summaries of all variables:
```{r}
airquality %>% select(-c(trtid10, STATEFP00, COUNTYFP00)) %>% 
  select(where(~is.numeric(.x) && !is.na(.x))) %>% 
  as.data.frame() %>% 
  stargazer(., type = "text", title = "Summary Statistics", digits = 0,
            summary.stat = c("mean", "sd", "min", "median", "max"))
```

## Visual Distributions

Visual representations of the data:

```{r, message=F, warning=F}
airquality %>% select(trtid10, percentile_1981, percentile_2016) %>% 
  pivot_longer(-trtid10, names_to = "measure", values_to = "value") %>% 
  ggplot(aes(x = value, fill = measure)) + 
  geom_histogram() + 
  facet_wrap(~measure, scales = "free") +
  xlim(0,100)
```

```{r}
airquality %>% 
  ggplot() +
  geom_point(aes(x=percentile_1981, y=percentile_2016)) +
  xlim(0, 100) +
  ylim(0, 100)
```

```{r}
airquality %>% 
  ggplot() +
  geom_point(aes(x=percentile_1981, y=pctile_change)) +
  xlim(0, 100) 
```

```{r}
airquality %>% 
  ggplot() +
  geom_point(aes(x=percentile_2016, y=pctile_change)) +
  xlim(0, 100) 
```


## Spatial Distributions

```{r, include = F}
cville_tracts <- readRDS("~/Desktop/Equity Center/summer-sandbox/cville_region_collection/data/cville_tracts.RDS")

airquality <- airquality %>% rename(GEOID = trtid10)

cvilleshapes <- merge(cville_tracts, airquality, by = 'GEOID', all.x = T)
cvilleshapes <- st_transform(cvilleshapes, crs = 4326) # to WGS84, given error
```

### Percentile in 1981
```{r}
pal <- colorNumeric("plasma", reverse = TRUE, domain = cvilleshapes$percentile_1981)
leaflet(cvilleshapes) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = cvilleshapes,
              fillColor = ~pal(percentile_1981),
              weight = 1,
              opacity = 1,
              color = "white",
              fillOpacity = 0.6,
              highlight = highlightOptions(weight = 2, fillOpacity = 0.8, bringToFront = T),
              popup = paste0("Tract Number: ", cvilleshapes$GEOID, "<br>",
                             "Percentile: ", round(cvilleshapes$percentile_1981, 2))) %>%
  addLegend("bottomright", pal = pal, values = cvilleshapes$percentile_1981,
            title = "Percentile, 1981", opacity = 0.7)
```

### Percentile in 2016
```{r}
pal <- colorNumeric("plasma", reverse = TRUE, domain = cvilleshapes$percentile_2016)
leaflet(cvilleshapes) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = cvilleshapes,
              fillColor = ~pal(percentile_2016),
              weight = 1,
              opacity = 1,
              color = "white",
              fillOpacity = 0.6,
              highlight = highlightOptions(weight = 2, fillOpacity = 0.8, bringToFront = T),
              popup = paste0("Tract Number: ", cvilleshapes$GEOID, "<br>",
                             "Percentile: ", round(cvilleshapes$percentile_2016, 2))) %>%
  addLegend("bottomright", pal = pal, values = cvilleshapes$percentile_2016,
            title = "Percentile, 2016", opacity = 0.7)
```


### Percentile Change, 1981-2016
```{r}
pal <- colorNumeric("plasma", reverse = TRUE, domain = cvilleshapes$pctile_change)
leaflet(cvilleshapes) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = cvilleshapes,
              fillColor = ~pal(pctile_change),
              weight = 1,
              opacity = 1,
              color = "white",
              fillOpacity = 0.6,
              highlight = highlightOptions(weight = 2, fillOpacity = 0.8, bringToFront = T),
              popup = paste0("Tract Number: ", cvilleshapes$GEOID, "<br>",
                             "Percentile Change: ", round(cvilleshapes$pctile_change, 2))) %>%
  addLegend("bottomright", pal = pal, values = cvilleshapes$pctile_change,
            title = "Percentile Change, 1981-2016", opacity = 0.7)
```

### PM~2.5~ Concentration in 2016
```{r}
pal <- colorNumeric("plasma", reverse = TRUE, domain = cvilleshapes$pm2_5_2016)
leaflet(cvilleshapes) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = cvilleshapes,
              fillColor = ~pal(pm2_5_2016),
              weight = 1,
              opacity = 1,
              color = "white",
              fillOpacity = 0.6,
              highlight = highlightOptions(weight = 2, fillOpacity = 0.8, bringToFront = T),
              popup = paste0("Tract Number: ", cvilleshapes$GEOID, "<br>",
                             "Concentration: ", cvilleshapes$pm2_5_2016)) %>%
  addLegend("bottomright", pal = pal, values = cvilleshapes$pm2_5_2016,
            title = "PM2.5 Concentration, 2016", opacity = 0.7)
```


### Change in PM~2.5~, 1981-2016
```{r}
pal <- colorNumeric("plasma", reverse = TRUE, domain = cvilleshapes$PM_change)
leaflet(cvilleshapes) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addPolygons(data = cvilleshapes,
              fillColor = ~pal(PM_change),
              weight = 1,
              opacity = 1,
              color = "white",
              fillOpacity = 0.6,
              highlight = highlightOptions(weight = 2, fillOpacity = 0.8, bringToFront = T),
              popup = paste0("Tract Number: ", cvilleshapes$GEOID, "<br>",
                             "PM2.5 Change: ", round(cvilleshapes$PM_change, 2))) %>%
  addLegend("bottomright", pal = pal, values = cvilleshapes$PM_change,
            title = "Change in PM2.5, 1981-2016", opacity = 0.7)
```


## Important Notes

The original data uses 2000 census tracts, since that is roughly the midpoint of their 1981-2016 time frame. Since we want to be able to look at this in relation to other tract-level data, we needed this in 2010 census tracts. To estimate the data for 2010 census tracts, I used the bridging data found [here](https://s4.ad.brown.edu/Projects/Diversity/Researcher/Bridging.htm).

Since the conversion to 2010 census tracts involved some estimation, the PM~2.5~ levels in the data may not be as accurate as the original 2000 data.