---
title: "Charlottesville CDC Places"
author: "Lee LeBoeuf"
date: "10/7/2021"
output: 
  html_document: 
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
invisible(lapply(list('tidyverse', 'stargazer', 'janitor', 'leaflet', 'rcartocolor', 'sf',
                      'RColorBrewer', 'viridis', 'rgdal', 'googlesheets4','stringr'),
                 function(pkg) library(pkg, character.only = TRUE)))

meta <- read_sheet("https://docs.google.com/spreadsheets/d/1nqm3DuVXD1ObbVe_deacvT7uSLdBXfQJo3mkbqDwrVo/edit?usp=sharing", sheet = "cdc_places", gs4_deauth())

# Main data
cvldat <- read.csv("../data/cdcplaces_cville_tract.csv")

# Spatial data
shape <- readRDS('../data/cville_tracts.RDS')
shape <- st_transform(shape, crs = 4326) # for leaflet
```

## Data Source

* These data come from the Center for Disease Control and Prevention (CDC) PLACES: Local Data for Better Health.
* More information about the CDC PLACES data can be found [here](https://www.cdc.gov/places/index.html). 
* Data shown here were downloaded using the CDC PLACES API in October of 2021. Documentation for the API can be found [here](https://dev.socrata.com/foundry/chronicdata.cdc.gov/cwsq-ngmh) and documentation for the data can be found [here](https://chronicdata.cdc.gov/500-Cities-Places/PLACES-Local-Data-for-Better-Health-Census-Tract-D/cwsq-ngmh). 
* The CDC has not released an update schedule for CDC PLACES data.

## The Data Used Here

* PLACES is a collaboration between the CDC, the Robert Wood Johnson Foundation, and the CDC Foundation. 
* The data provide model-based estimates for chronic disease risk factors, health outcomes, and clinical prevention services. According the the website, "CDC uses an innovative peer-reviewed multilevel regression and poststratification (MPR) approach that links geocoded health surveys and high spatial resolution population demographic and socioeconomic data."
* CDC pulls data from the CDC Behavioral Risk Factor Surveillance System, the Census 2010 population, and the American Community Survey estimates.
* The full data from the CDC include 27 measures: 5 unhealthy behaviors, 13 health outcomes, and 9 prevention practices. 
* The data used here include a total of 17 variables. Below, we show a number of them grouped into health outcomes, health behaviors and health access. 
* All of the data values represent the predicted percent prevalence of a given outcome within a census tract. For almost all measures, respondents to health surveys answered whether they had ever been diagnosed with a given outcome by a health care professional. For many indicators, the estimated prevalance is likely to be an underestimate, particularly in areas where residents have limited access to health care and are therefore less likely to receive a medical diagnosis for a particular ailment. 
* Data for some outcomes are from 2017, but most are from 2018. 
* Data are available at the tract levels for Albemarle, Charlottesville, Louisa, Greene, Fluvanna, Nelson counties.

## Variable descriptions
```{r}
#meta %>% 
#  filter(su_tract == 1) %>%
#  select(varname, about) %>% as.list()

glimpse(cvldat)

cvldat %>% select(-c(locationname, countyname, totalpopulation)) %>% 
  select(where(~is.numeric(.x))) %>% 
  as.data.frame() %>% 
  stargazer(., type = "text", title = "Summary Statistics", digits = 2,
            summary.stat = c("mean", "sd", "min", "median", "max"))
```

## Visual distribution 
```{r}
longdat <- cvldat %>% select(-c(countyname, totalpopulation)) %>% 
  pivot_longer(-locationname, names_to = "measure", values_to = "value") 


longdat$measure <- factor(longdat$measure,
                         levels = c("Coronary_Heart_Disease2018", "Binge_Drinking2018", "Mental_Health2018",
                                    "High_Blood_Pressure2017", "Physical_Inactivity2018", "Diabetes2018",
                                    "Current_Smoking2018", "Cancer_except_skin2018", "Current_Asthma2018",
                                    "Dental_Visit2018", "High_Cholesterol2017", "COPD2018", "Obesity2018",
                                    "Physical_Health2018", "Health_Insurance2018", "less_than_sevenhr_sleep2018",
                                    "Annual_Checkup2018"))

longdat %>%
  ggplot(aes(x = value, fill = measure)) +
  scale_fill_viridis(option = "Blues", discrete = TRUE, guide = FALSE) +
  geom_histogram() + 
  facet_wrap(~measure, scales = "free", ncol = 4)
```

```{r, results = "asis"}
meta %>% 
  filter(varname %in% c("Coronary_Heart_Disease2018", "Binge_Drinking2018", "Mental_Health2018",
                        "High_Blood_Pressure2017", "Physical_Inactivity2018", "Diabetes2018",
                        "Current_Smoking2018",
                        "Cancer_except_skin2018", "Current_Asthma2018", "Dental_Visit2018",
                        "High_Cholesterol2017", "COPD2018", "Obesity2018", "Physical_Health2018",
                        "Health_Insurance2018", "less_than_sevenhr_sleep2018", "Annual_Checkup2018")) %>%
  mutate(label = paste0(varname, ": ", about)) %>% 
  select(label) %>% 
  as.list()
```

## Health outcomes {.tabset}
```{r, echo = FALSE}
cvldat <- cvldat %>%
  rename(GEOID = locationname) %>% 
   mutate(GEOID = as.character(GEOID))

mapdat <- shape %>% 
  left_join(cvldat, by = "GEOID")
```

### Asthma
* Adjusted percent of survey respondents aged >= 18 who reported that they had ever been told by a health professional that they have asthma __and__ that they still have asthma.
```{r}
pal <- colorNumeric("Blues", domain = mapdat$Current_Asthma2018)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat,
              fillColor = ~pal(mapdat$Current_Asthma2018),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T),
              popup = paste0("GEOID: ", mapdat$GEOID, "<br>",
                             "Percent of adults with asthma: ", mapdat$Current_Asthma2018)) %>% 
  leaflet::addLegend("bottomright", pal = pal, values = (mapdat$Current_Asthma2018), 
            title = "Predicted % of adults <br> with current asthma <br> in 2018", opacity = 0.7)

```

### Coronary Heart Disease
* Adjusted percent of survey respondents aged >= 18 who reported that they had ever been told by a health professional that they had angina or coronary hearth disease. 
```{r}
pal <- colorNumeric("Blues", domain = mapdat$Coronary_Heart_Disease2018)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat,
              fillColor = ~pal(mapdat$Coronary_Heart_Disease2018),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T),
              popup = paste0("GEOID: ", mapdat$GEOID, "<br>",
                             "Percent of adults with Coronary Heart Disease: ", mapdat$Coronary_Heart_Disease2018)) %>% 
  leaflet::addLegend("bottomright", pal = pal, values = (mapdat$Coronary_Heart_Disease2018), 
            title = "Predicted % of adults <br> with Coronary Heart Disease <br> in 2018", opacity = 0.7)

```

### Cancer (excluding skin cancer)
* Adjusted percent of survey respondents aged >= 18 who reported that they had ever been told by a health professional that they have any type of cancer except skin cancer. This variable is not specific to any type of cancer.
```{r}
pal <- colorNumeric("Blues", domain = mapdat$Cancer_except_skin2018)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat,
              fillColor = ~pal(mapdat$Cancer_except_skin2018),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T),
              popup = paste0("GEOID: ", mapdat$GEOID, "<br>",
                             "Percent of adults with cancer: ", mapdat$Coronary_Heart_Disease2018)) %>% 
  leaflet::addLegend("bottomright", pal = pal, values = (mapdat$Cancer_except_skin2018), 
            title = "Predicted % of adults <br> with cancer (excluding skin) <br> in 2018", opacity = 0.7)

```

### Diabetes 
* Adjusted percent of survey respondents aged >= 18 who reported that they had ever been told by a health professional that they had diabetes (excluding diabetes during pregnancy). 
```{r}
pal <- colorNumeric("Blues", domain = mapdat$Diabetes2018)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat,
              fillColor = ~pal(mapdat$Diabetes2018),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T),
              popup = paste0("GEOID: ", mapdat$GEOID, "<br>",
                             "Percent of adults with diabetes: ", mapdat$Diabetes2018)) %>% 
  leaflet::addLegend("bottomright", pal = pal, values = (mapdat$Diabetes2018), 
            title = "Predicted % of adults <br> with diabetes <br> in 2018", opacity = 0.7)

```

### Obesity
* Adjusted percent of survey respondents aged >= 18 who reported that they have a BMI >= 30kg/m^2. This indicator is calculated by the CDC from self-reported weight and height, excluding pregnant women. Self-reported height and weight tend to lead to lower BMI estimates than clinical height and weight measurements. 
```{r}
pal <- colorNumeric("Blues", domain = mapdat$Obesity2018)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat,
              fillColor = ~pal(mapdat$Obesity2018),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T),
              popup = paste0("GEOID: ", mapdat$GEOID, "<br>",
                             "Percent of adults who have a BMI >= 30: ", mapdat$Obesity2018)) %>% 
  leaflet::addLegend("bottomright", pal = pal, values = (mapdat$Obesity2018), 
            title = "Predicted % of adults <br> who had a BMI >= 30 <br> in 2018", opacity = 0.7)

```

## Health access {.tabset}

### Lack of health insurance 
* Adjusted percent of respondents aged 18-64 who reported have no current health insurance. 
```{r}
pal <- colorNumeric("Blues", domain = mapdat$Health_Insurance2018)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat,
              fillColor = ~pal(mapdat$Health_Insurance2018),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T),
              popup = paste0("GEOID: ", mapdat$GEOID, "<br>",
                             "Percent of adults without health insurance: ", mapdat$Health_Insurance2018)) %>% 
  leaflet::addLegend("bottomright", pal = pal, values = (mapdat$Health_Insurance2018), 
            title = "Predicted % of adults <br> without health insurance <br> in 2018", opacity = 0.7)

```

### Routine check-up in the last year in 2018 
* Adjusted percent of survey respondents aged >= 18 who reported having been to the doctor for a routine checkup (excluding visits for a specific injury or illness) in the previous year. 
```{r}
pal <- colorNumeric("Blues", domain = mapdat$Annual_Checkup2018)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat,
              fillColor = ~pal(mapdat$Annual_Checkup2018),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T),
              popup = paste0("GEOID: ", mapdat$GEOID, "<br>",
                             "Percent of adults who went <br> to the doctor in last year: ", mapdat$Annual_Checkup2018)) %>% 
  leaflet::addLegend("bottomright", pal = pal, values = (mapdat$Annual_Checkup2018), 
            title = "Predicted % of adults <br> who went to the doctor <br> in the last year <br> in 2018", opacity = 0.7)

```

### Visits to dentist in the last year in 2018
* Adjusted percent of survey respondents aged >= 18 who reported having been to a dentist or dental clinic in the previous year. 
```{r}
pal <- colorNumeric("Blues", domain = mapdat$Dental_Visit2018)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat,
              fillColor = ~pal(mapdat$Dental_Visit2018),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T),
              popup = paste0("GEOID: ", mapdat$GEOID, "<br>",
                             "Percent of adults who went <br> to the dentist in last year: ", mapdat$Dental_Visit2018)) %>% 
  leaflet::addLegend("bottomright", pal = pal, values = (mapdat$Dental_Visit2018), 
            title = "Predicted % of adults <br> who went to the dentist <br> in the last year <br> in 2018", opacity = 0.7)

```

## Health behaviors {.tabset}

### Curent smoking in 2018 
* Adjusted percent of survey respondents aged >= 18 who reported having smoked >= 100 cigarettes in their lifetime and currently smoke every day or some days. This indicator is not a measure of intent or attempts to quit smoking or exposure to secondhand smoke among non-smokers. 
```{r}
pal <- colorNumeric("Blues", domain = mapdat$Current_Smoking2018)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat,
              fillColor = ~pal(mapdat$Current_Smoking2018),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T),
              popup = paste0("GEOID: ", mapdat$GEOID, "<br>",
                             "Percent of adults who smoke: ", mapdat$Current_Smoking2018)) %>% 
  leaflet::addLegend("bottomright", pal = pal, values = (mapdat$Current_Smoking2018), 
            title = "Predicted % of adults <br> who currently smoke <br> in 2018", opacity = 0.7)

```

### Less than 7 hours of sleep
* Adjusted percent of survey respondents aged >= 18 who reported usually getting insufficient sleep (<7 hours for those aged >=18 years, during a 24 hour period). This indicator does not measure variation in sleep duration, sleep quality, or specific sleep problems. 
```{r}
pal <- colorNumeric("Blues", domain = mapdat$less_than_sevenhr_sleep2018)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat,
              fillColor = ~pal(mapdat$less_than_sevenhr_sleep2018),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T),
              popup = paste0("GEOID: ", mapdat$GEOID, "<br>",
                             "Percent of adults who average <7 hrs <br> of sleep a night in 2018: ", mapdat$less_than_sevenhr_sleep2018)) %>% 
  leaflet::addLegend("bottomright", pal = pal, values = (mapdat$less_than_sevenhr_sleep2018), 
            title = "Predicted % of adults <br> who average <7 hrs <br> of sleep a night in 2018", opacity = 0.7)

```


