---
title: "DAYMET Temperature and Precipitation Data"
author: "Tolu Odukoya & Lee LeBoeuf"
date: "9/27/2021"
output: 
  html_document:
    toc: yes
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
invisible(lapply(list('tidyverse', 'stargazer', 'janitor', 'tigris', 'sf', 'leaflet', 'rcartocolor', 
                      'RColorBrewer', 'viridis', 'rgdal', 'lodes', 'psych', 'reshape2', 'googlesheets4', 
                      'sp', 'geosphere', 'stringr', 'ggrepel', 'transformr', 'gifski', 'dygraphs', 'xts'),
                 function(pkg) library(pkg, character.only = TRUE)))
meta <- read_sheet("https://docs.google.com/spreadsheets/d/1nqm3DuVXD1ObbVe_deacvT7uSLdBXfQJo3mkbqDwrVo/edit?usp=sharing", sheet = "DAYMET", gs4_deauth())
cvldat <- read.csv("daymet_cville_county.csv")
cvldatblkgr <- read.csv("daymet_cville_blkgrps.csv")
```

## Data Source

* These data come from Daymet: Monthly Climate Summaries on a 1-km Grid for North America.
* More information about the DAYMET data can be found [here](https://daac.ornl.gov/cgi-bin/dsviewer.pl?ds_id=1855). 
* Data shown here were downloaded using the development version of the FedData package in R in September of 2021. Documentation can be found [here](https://github.com/ropensci/FedData). 

## The Data Used Here

* Daymet is a data product derived from a collection of algorithms and computer software designed to interpolate and extrapolate from daily meteorological observations to produce gridded estimates of daily weather parameters. Weather parameters generated include daily surfaces of minimum and maximum temperature, precipitation, vapor pressure, radiation, snow water equivalent, and day length produced on a 1 km x 1 km gridded surface. 
* The motivation for producing Daymet is to provide measurements of near-surface meteorological conditions where no instrumentation exists. Having estimates of these surfaces is critical to understanding many processes in the terrestrial biogeochemical system.
* The data shown here includes the monthly average maximum temperature for the months June, July, and August, the average maximum temperature across the entire year, and the annual total precipitation. The data covers years from 1980-2020. The maximum temperature was originally recorded in Celsuis and the precipitation in Millimeters, they were both converted to Fahrenheit and Inches respectively. 
* Data are available at the block, block groups, tract, and county levels for Albemarle, Charlottesville, Louisa, Greene, Fluvanna, Nelson.
* To Learn More, See: [Daily Surface Weather and Climatological Summaries](https://daymet.ornl.gov/overview)
* Below we show summaries of the data at the block group level, and then we show variation by county over time. 

## Variable descriptions
These summaries are based on data at the block group level from the years 1980-2020. Data are also available at the block, tract, and county levels. 
```{r}
meta %>% 
  filter(su_blkgp == 1) %>%
  select(varname, about) %>% as.list()
glimpse(cvldat)
cvldat %>% select(June_AvgMaxTF, July_AvgMaxTF, Aug_AvgMaxTF, AvgAnnMaxTF, TotpercInch) %>% 
  select(where(~is.numeric(.x))) %>% 
  as.data.frame() %>% 
  stargazer(., type = "text", title = "Summary Statistics", digits = 2,
            summary.stat = c("mean", "sd", "min", "median", "max"))
```

## Visual distribution 
These distributions show data at the block group level from the years 1980-2020. Data are also available at the block, tract, and county levels. 
```{r}
longdat <- cvldat %>% select(c(GEOID, June_AvgMaxTF, July_AvgMaxTF, Aug_AvgMaxTF, AvgAnnMaxTF, TotpercInch)) %>% pivot_longer(-GEOID, names_to = "measure", values_to = "value") 
longdat$measure <- factor(longdat$measure,
                         levels = c("June_AvgMaxTF", "July_AvgMaxTF", "Aug_AvgMaxTF", "AvgAnnMaxTF", "TotpercInch"))
longdat %>%
  ggplot(aes(x = value, fill = measure)) +
  scale_fill_viridis(option = "plasma", discrete = TRUE, guide = FALSE) +
  geom_histogram() + 
  facet_wrap(~measure, scales = "free")
```

```{r, results = "asis"}
meta %>% 
  filter(varname %in% c("June_AvgMaxTF", "July_AvgMaxTF", "Aug_AvgMaxTF", "AvgAnnMaxTF", "TotpercInch")) %>%
  mutate(label = paste0(varname, ": ", about)) %>% 
  select(label) %>% 
  as.list()
```

## Block Group Variation in 2020 {.tabset}
All of the following maps show data from the year 2020, but data are available for years 1980-2020. 
```{r, echo = FALSE}
shape <- readRDS('cville_blkgps.RDS')
shape <- st_transform(shape, crs = 4326) # to WGS84, given error
mapdat <- merge(shape, cvldatblkgr, by = 'GEOID', all.x = T)
```

### Average maximum annual temperature {.tabset}
```{r}
mapdat2020 <- mapdat[which(mapdat$year == 2020),]

pal <- colorNumeric("plasma", reverse = TRUE, domain = mapdat2020$AvgAnnMaxTF)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat2020,
              fillColor = ~pal(mapdat2020$AvgAnnMaxTF),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T),
              popup = paste0("GEOID: ", mapdat2020$GEOID, "<br>",
                             "Temperature in Fahrenhiet: ", mapdat2020$AvgAnnMaxTF)) %>% 
  leaflet::addLegend("bottomright", pal = pal, values = (mapdat2020$AvgAnnMaxTF), 
            title = "Average annual <br> temperature <br> in Fahrenheit <br> in 2020", opacity = 0.7)

```

### Average maximum August temperature
Monhtly summaries are also available for June and July. 
```{r}
pal <- colorNumeric("plasma", reverse = TRUE, domain = mapdat2020$Aug_AvgMaxTF)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat2020,
              fillColor = ~pal(mapdat2020$Aug_AvgMaxTF),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T),
              popup = paste0("GEOID: ", mapdat2020$GEOID, "<br>",
                             "Temperature in Fahrenhiet: ", mapdat2020$Aug_AvgMaxTF)) %>% 
  leaflet::addLegend("bottomright", pal = pal, values = (mapdat2020$Aug_AvgMaxTF), 
            title = "Average August max <br> temperature <br> in Fahrenheit <br> in 2020", opacity = 0.7)

```

### Annual precipitation
```{r}
pal <- colorNumeric("plasma", reverse = TRUE, domain = mapdat2020$TotpercInch)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat2020,
              fillColor = ~pal(mapdat2020$TotpercInch),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("GEOID: ", mapdat2020$GEOID, "<br>",
                             "Precipitation in Inches: ", mapdat2020$TotpercInch)
              ) %>% 
  leaflet::addLegend("bottomright", pal = pal, values = (mapdat2020$TotpercInch), 
            title = "Annual precipitation <br>in Inches <br>in 2020", opacity = 0.7)

```

## County variation over time {.tabset}

### Annual average temperature {.tabset}
```{r, echo=FALSE}
cville_sf1p <- cvldat %>% dplyr::select(NAMELSAD, year, AvgAnnMaxTF)
cville_sf1g <- filter(cville_sf1p, NAMELSAD == "Greene County")
cville_sf1al <- filter(cville_sf1p, NAMELSAD == "Albemarle County") %>%
  rename(NAMELSAD1 = NAMELSAD, AvgAnnMaxTF1 = AvgAnnMaxTF, year1 = year)
cville_sf1c<- filter(cville_sf1p, NAMELSAD == "Charlottesville city") %>%
  rename(NAMELSAD2 = NAMELSAD, AvgAnnMaxTF2 = AvgAnnMaxTF, year2 = year)
cville_sf1f<- filter(cville_sf1p, NAMELSAD == "Fluvanna County") %>%
  rename(NAMELSAD3 = NAMELSAD, AvgAnnMaxTF3 = AvgAnnMaxTF, year3 = year)
cville_sf1l <- filter(cville_sf1p, NAMELSAD == "Louisa County") %>%
  rename(NAMELSAD4 = NAMELSAD, AvgAnnMaxTF4 = AvgAnnMaxTF, year4 = year)
cville_sf1n <- filter(cville_sf1p, NAMELSAD == "Nelson County") %>%
  rename(NAMELSAD5 = NAMELSAD, AvgAnnMaxTF5 = AvgAnnMaxTF, year5 = year)

cville_sf1x <- cbind(cville_sf1g, cville_sf1al, cville_sf1c, cville_sf1f, cville_sf1l, cville_sf1n) %>%
  rename(Greene = AvgAnnMaxTF, Albemarle = AvgAnnMaxTF1, Charlottesville = AvgAnnMaxTF2, Fluvanna = AvgAnnMaxTF3, Louisa = AvgAnnMaxTF4, Nelson = AvgAnnMaxTF5)

cville_sf1x <- cville_sf1x[ , -which(names(cville_sf1x) %in% c("year1","year2", "year3", "year4", "year5", "NAMELSAD", "NAMELSAD1", "NAMELSAD2", "NAMELSAD3", "NAMELSAD4","NAMELSAD5"))]
```

```{r}
dygraph(cville_sf1x, main = "Average maximum annual temperature for each county over time", ylab = "Temperature (F)") %>%
  dySeries("Greene", label = "Greene") %>%
  dySeries("Albemarle", label = "Albemarle") %>%
  dySeries("Charlottesville", label = "Charlottesville") %>%
  dySeries("Fluvanna", label = "Fluvanna") %>%
  dySeries("Louisa", label = "Louisa") %>%
  dySeries("Nelson", label = "Nelson") %>%
  dyHighlight(highlightCircleSize = 3, 
              highlightSeriesBackgroundAlpha = 0.2,
              hideOnMouseOut = FALSE, highlightSeriesOpts = list(strokeWidth = 3)) %>% 
dyLegend(show = "always", hideOnMouseOut = FALSE, width = 400) %>% 
    dyRangeSelector(height = 20) %>%  dyOptions(colors = RColorBrewer::brewer.pal(6, "Set2"))
```

### Average maximum August temperatures 
```{r, echo=FALSE}
cville_sf1p <- cvldat %>% dplyr::select(NAMELSAD, year, Aug_AvgMaxTF)
cville_sf1g <- filter(cville_sf1p, NAMELSAD == "Greene County")
cville_sf1al <- filter(cville_sf1p, NAMELSAD == "Albemarle County") %>%
  rename(NAMELSAD1 = NAMELSAD, Aug_AvgMaxTF1 = Aug_AvgMaxTF, year1 = year)
cville_sf1c<- filter(cville_sf1p, NAMELSAD == "Charlottesville city") %>%
  rename(NAMELSAD2 = NAMELSAD, Aug_AvgMaxTF2 = Aug_AvgMaxTF, year2 = year)
cville_sf1f<- filter(cville_sf1p, NAMELSAD == "Fluvanna County") %>%
  rename(NAMELSAD3 = NAMELSAD, Aug_AvgMaxTF3 = Aug_AvgMaxTF, year3 = year)
cville_sf1l <- filter(cville_sf1p, NAMELSAD == "Louisa County") %>%
  rename(NAMELSAD4 = NAMELSAD, Aug_AvgMaxTF4 = Aug_AvgMaxTF, year4 = year)
cville_sf1n <- filter(cville_sf1p, NAMELSAD == "Nelson County") %>%
  rename(NAMELSAD5 = NAMELSAD, Aug_AvgMaxTF5 = Aug_AvgMaxTF, year5 = year)

cville_sf1x <- cbind(cville_sf1g, cville_sf1al, cville_sf1c, cville_sf1f, cville_sf1l, cville_sf1n) %>%
  rename(Greene = Aug_AvgMaxTF, Albemarle = Aug_AvgMaxTF1, Charlottesville = Aug_AvgMaxTF2, Fluvanna = Aug_AvgMaxTF3, Louisa = Aug_AvgMaxTF4, Nelson = Aug_AvgMaxTF5)

cville_sf1x <- cville_sf1x[ , -which(names(cville_sf1x) %in% c("year1","year2", "year3", "year4", "year5", "NAMELSAD", "NAMELSAD1", "NAMELSAD2", "NAMELSAD3", "NAMELSAD4","NAMELSAD5"))]
```

```{r}
dygraph(cville_sf1x, main = "Average maximum August temperature for each county over time", ylab = "Temperature (F)") %>%
  dySeries("Greene", label = "Greene") %>%
  dySeries("Albemarle", label = "Albemarle") %>%
  dySeries("Charlottesville", label = "Charlottesville") %>%
  dySeries("Fluvanna", label = "Fluvanna") %>%
  dySeries("Louisa", label = "Louisa") %>%
  dySeries("Nelson", label = "Nelson") %>%
  dyHighlight(highlightCircleSize = 3, 
              highlightSeriesBackgroundAlpha = 0.2,
              hideOnMouseOut = FALSE, highlightSeriesOpts = list(strokeWidth = 3)) %>% 
dyLegend(show = "always", hideOnMouseOut = FALSE, width = 400) %>% 
    dyRangeSelector(height = 20) %>%  dyOptions(colors = RColorBrewer::brewer.pal(6, "Set2"))
```

### Annual precipitation 
```{r, echo=FALSE}
cville_sf1p <- cvldat %>% dplyr::select(NAMELSAD, year, TotpercInch)
cville_sf1g <- filter(cville_sf1p, NAMELSAD == "Greene County")
cville_sf1al <- filter(cville_sf1p, NAMELSAD == "Albemarle County") %>%
  rename(NAMELSAD1 = NAMELSAD, TotpercInch1 = TotpercInch, year1 = year)
cville_sf1c<- filter(cville_sf1p, NAMELSAD == "Charlottesville city") %>%
  rename(NAMELSAD2 = NAMELSAD, TotpercInch2 = TotpercInch, year2 = year)
cville_sf1f<- filter(cville_sf1p, NAMELSAD == "Fluvanna County") %>%
  rename(NAMELSAD3 = NAMELSAD, TotpercInch3 = TotpercInch, year3 = year)
cville_sf1l <- filter(cville_sf1p, NAMELSAD == "Louisa County") %>%
  rename(NAMELSAD4 = NAMELSAD, TotpercInch4 = TotpercInch, year4 = year)
cville_sf1n <- filter(cville_sf1p, NAMELSAD == "Nelson County") %>%
  rename(NAMELSAD5 = NAMELSAD, TotpercInch5 = TotpercInch, year5 = year)

cville_sf1x <- cbind(cville_sf1g, cville_sf1al, cville_sf1c, cville_sf1f, cville_sf1l, cville_sf1n) %>%
  rename(Greene = TotpercInch, Albemarle = TotpercInch1, Charlottesville = TotpercInch2, Fluvanna = TotpercInch3, Louisa = TotpercInch4, Nelson = TotpercInch5)

cville_sf1x <- cville_sf1x[ , -which(names(cville_sf1x) %in% c("year1","year2", "year3", "year4", "year5", "NAMELSAD", "NAMELSAD1", "NAMELSAD2", "NAMELSAD3", "NAMELSAD4","NAMELSAD5"))]
```

```{r}
dygraph(cville_sf1x, main = "Annual precipitation for each county over time", ylab = "Precipitation (In)") %>%
  dySeries("Greene", label = "Greene") %>%
  dySeries("Albemarle", label = "Albemarle") %>%
  dySeries("Charlottesville", label = "Charlottesville") %>%
  dySeries("Fluvanna", label = "Fluvanna") %>%
  dySeries("Louisa", label = "Louisa") %>%
  dySeries("Nelson", label = "Nelson") %>%
  dyHighlight(highlightCircleSize = 3, 
              highlightSeriesBackgroundAlpha = 0.2,
              hideOnMouseOut = FALSE, highlightSeriesOpts = list(strokeWidth = 3)) %>% 
dyLegend(show = "always", hideOnMouseOut = FALSE, width = 400) %>% 
    dyRangeSelector(height = 20) %>%  dyOptions(colors = RColorBrewer::brewer.pal(6, "Set2"))
```




