---
title: "Charlottesville Region Broadband Availability"
author: "Lee LeBoeuf, Michele Claibourn"
date: "7/30/2021"
output: 
  html_document: 
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

invisible(lapply(list('tidyverse', 'stargazer', 'janitor', 'tigris', 'sf', 'leaflet', 'rcartocolor', 
                      'RColorBrewer', 'viridis', 'googlesheets4', 'stringr'),
                 function(pkg) library(pkg, character.only = TRUE)))

meta <- read_sheet("https://docs.google.com/spreadsheets/d/1nqm3DuVXD1ObbVe_deacvT7uSLdBXfQJo3mkbqDwrVo/edit?usp=sharing", sheet = "fcc_broadband", gs4_deauth())

# main data
cvltracts <- read.csv("../data/fcc_broadband_cville_tract.csv")
cvltracts <- cvltracts %>% 
  mutate(tract = as.character(tract))


cvlblkgps <- read.csv("../data/fcc_broadband_cville_blkgr.csv")
cvlblkgps <- cvlblkgps %>% 
  mutate(Blkgr = as.character(Blkgr))

cvlblocks <- read.csv("../data/fcc_broadband_cville_block.csv")
cvlblocks <- cvlblocks %>% 
  mutate(BlockCode = as.character(BlockCode))

# spatial data
shape_tracts <- readRDS('../data/cville_tracts.RDS')
shape_tracts <- st_transform(shape_tracts, crs = 4326) # to WGS84, given error

shape_blkgps <- readRDS('../data/cville_blkgps.RDS')
shape_blkgps <- st_transform(shape_blkgps, crs = 4326) # to WGS84, given error

shape_blocks <- readRDS('../data/cville_blocks.RDS')
shape_blocks <- st_transform(shape_blocks, crs = 4326) # to WGS84, given error
```

## Data Source
* These data come from the FCC's Form 477 Fixed Broadband Deployment Data. 
* FCC data is released biannually on a year and half delay from when it was collected. New data can be downloaded from this [link](https://www.fcc.gov/general/broadband-deployment-data-fcc-form-477). 
* Data presented here are from June of 2020. 

## Data specifics 
* All internet providers must submit form 477 to the FCC biannually detailing the Maximum download and upload speeds (Mbps) they advertise to consumers along with other information about the company. 
* The FCC makes these data publicly available on a year and a half delay from when all 477 forms are submitted. 
    + Limitations: (1) The data are available at the census block level, but are likely to overstate coverage; if an internet provider provides internet to a single person in a census block, the entire census block is marked as having coverage by that provider. 
    + Strengths: (1) These are the same data used by the government to make policy decisions regarding broadband access. (2) The [Broadband Deployment Accuracy and Technological Availability (DATA) Act](https://www.commerce.senate.gov/2020/3/bill-to-improve-broadband-data-maps-signed-into-law) of March 2020 is intended to reduce the limitation listed above. 
* The FCC has established benchmarks for "advanced telecommunications capability", services that enables "users to originate and receive high-quality voice, data, graphics, and video telecommunications." The current benchmarks are set at speeds of 25 (download) and 3 (upload) Mbps ([2020 Broadband Deployment Report](https://docs.fcc.gov/public/attachments/FCC-20-50A1.pdf)).
    
## Variable descriptions

Variables include: 

```{r}
# meta %>% 
#   filter(su_blkgp == 1) %>%
#   select(varname, about) %>% as.list()

glimpse(cvlblkgps)
```

* BlockCode: geographic identifer for spatial area (tracts, block groups, and blocks)
* Blkgr: geographic identifer for spatial area (tracts, block groups, and blocks)
* tract: geographic identifer for spatial area (tracts, block groups, and blocks)

* resproviders: The number of residential internet providers in the block/block group/tract
* bb253_num: The number of residential broadband providers providing service that meets the FCC benchmark for "advanced telecommunications capability", 25/3 Mbps 
* bb253_per: The percent of residential broadband providers providing service that meets the FCC benchmark for "advanced telecommunications capability", 25/3 Mbps 
* bbmin_dl: The minimum advertised download speed provider in the block/block group/tract
* bbmax_dl: The maximum advertised download speed provider in the block/block group/tract
* bbmin_up: The minimum advertised upload speed provider in the block/block group/tract
* bbmax_up: The maximum advertised upload speed provider in the block/block group/tract
* avgMaxAdDown:The average maximum advertised download speed by each broadband provider in the block/block group/tract
* avgMaxAdUp: The average maximum advertised upload speed by each broadband provider in the block/block group/tract

5-number summaries of variables by block groups:

```{r}
cvlblkgps %>% select(-c(Blkgr, tract)) %>% 
  select(where(~is.numeric(.x))) %>% 
  as.data.frame() %>% 
  stargazer(., type = "text", title = "Summary Statistics", digits = 2,
            summary.stat = c("mean", "sd", "min", "median", "max"))
```

## Visual distribution {.tabset}

### Tracts
```{r}
cvltracts %>% 
  pivot_longer(-tract, names_to = "measure", values_to = "value") %>%
  mutate(measure = factor(measure, levels = c("resproviders", "bb253_num", "bb253_per",
                                              "avgMaxAdDown", "avgMaxAdUp",
                                              "bbmax_dl", "bbmax_up", "bbmin_dl", "bbmin_up"))) %>% 
  ggplot(aes(x = value, fill = measure)) +
  scale_fill_viridis(option = "plasma", discrete = TRUE, guide = "none") +
  geom_histogram() + 
  facet_wrap(~measure, scales = "free")
```

### Block groups
```{r}
cvlblkgps %>% select(-tract) %>% 
  pivot_longer(-Blkgr, names_to = "measure", values_to = "value") %>%
  mutate(measure = factor(measure, levels = c("resproviders", "bb253_num", "bb253_per",
                                              "avgMaxAdDown", "avgMaxAdUp",
                                              "bbmax_dl", "bbmax_up", "bbmin_dl", "bbmin_up"))) %>% 
  ggplot(aes(x = value, fill = measure)) +
  scale_fill_viridis(option = "plasma", discrete = TRUE, guide = "none") +
  geom_histogram() + 
  facet_wrap(~measure, scales = "free")
```

### Blocks
```{r}
cvlblocks %>% select(-tract, Blkgr) %>% 
  pivot_longer(-BlockCode, names_to = "measure", values_to = "value") %>%
  mutate(measure = factor(measure, levels = c("resproviders", "bb253_num", "bb253_per",
                                              "avgMaxAdDown", "avgMaxAdUp",
                                              "bbmax_dl", "bbmax_up", "bbmin_dl", "bbmin_up"))) %>% 
  ggplot(aes(x = value, fill = measure)) +
  scale_fill_viridis(option = "plasma", discrete = TRUE, guide = "none") +
  geom_histogram() + 
  facet_wrap(~measure, scales = "free")
```


## Mapping the data
```{r, echo = FALSE}
maptracts <- shape_tracts %>% 
  left_join(cvltracts, by = c("GEOID" = "tract"))

mapblkgps <- shape_blkgps %>% 
  left_join(cvlblkgps, by = c("GEOID" = "Blkgr"))
```
  
### Number of residential broadband providers{.tabset} 

#### Tracts
```{r }
pal <- colorNumeric("plasma", reverse = TRUE, domain = maptracts$resproviders)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = maptracts,
              fillColor = ~pal(resproviders),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("GEOID: ", maptracts$GEOID, "<br>",
                             "Number of residential providers: ", maptracts$resproviders)
  ) %>% 
  addLegend("bottomright", pal = pal, values = maptracts$resproviders, 
            title = "Number of residential <br> providers", opacity = 0.7)
```

#### Block groups
```{r }
pal <- colorNumeric("plasma", reverse = TRUE, domain = mapblkgps$resproviders)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapblkgps,
              fillColor = ~pal(resproviders),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("GEOID: ", mapblkgps$GEOID, "<br>",
                             "Number of residential providers: ", mapblkgps$resproviders)
  ) %>% 
  addLegend("bottomright", pal = pal, values = mapblkgps$resproviders, 
            title = "Number of residential <br> providers", opacity = 0.7)
```

#### Blocks

Block level maps are slow to render and may be added later.

### Average maximum advertised download speeds {.tabset}

#### Tracts
```{r }
pal <- colorNumeric("plasma", reverse = TRUE, domain = maptracts$avgMaxAdDown) 

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = maptracts,
              fillColor = ~pal(avgMaxAdDown),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("GEOID: ", maptracts$GEOID, "<br>",
                             "Average max advertised <br> download speeds: ", round(maptracts$avgMaxAdDown, 2))
  ) %>% 
  addLegend("bottomright", pal = pal, values = maptracts$avgMaxAdDown, 
            title = "Average maximum <br>advertised  <br>download speeds", opacity = 0.7)
```

#### Block groups
```{r }
pal <- colorNumeric("plasma", reverse = TRUE, domain = mapblkgps$avgMaxAdDown) 

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapblkgps,
              fillColor = ~pal(avgMaxAdDown),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("GEOID: ", mapblkgps$GEOID, "<br>",
                             "Average max advertised <br> download speeds: ", round(mapblkgps$avgMaxAdDown, 2))
  ) %>% 
  addLegend("bottomright", pal = pal, values = mapblkgps$avgMaxAdDown, 
            title = "Average maximum <br>advertised  <br>download speeds", opacity = 0.7)
```

#### Blocks

Block level maps are slow to render and may be added later.

### Maximum available advertised download speeds {.tabset}

#### Tracts
```{r }
pal <- colorNumeric("plasma", reverse = TRUE, domain = maptracts$bbmax_dl)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = maptracts,
              fillColor = ~pal(bbmax_dl),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("GEOID: ", maptracts$GEOID, "<br>",
                             "Max available <br> advertised download speeds: ", round(maptracts$bbmax_dl, 2))
  ) %>% 
  addLegend("bottomright", pal = pal, values = maptracts$bbmax_dl, 
            title = "Max available <br>advertised <br> download speeds", opacity = 0.7)
```

#### Block groups
```{r }
pal <- colorNumeric("plasma", reverse = TRUE, domain = mapblkgps$bbmax_dl)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapblkgps,
              fillColor = ~pal(bbmax_dl),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("GEOID: ", mapblkgps$GEOID, "<br>",
                             "Max available <br> advertised download speeds: ", round(mapblkgps$bbmax_dl, 2))
  ) %>% 
  addLegend("bottomright", pal = pal, values = mapblkgps$bbmax_dl, 
            title = "Max available <br>advertised <br> download speeds", opacity = 0.7)
```

#### Blocks

Block level maps are slow to render and may be added later.