---
title: "FEMA National Risk Index"
subtitle: "Data for Charlottesville Region"
author: "Michele Claibourn"
date: "6/24/2021"
output: 
  html_document: 
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# 0. Load libraries and data
library(tidyverse)
library(stargazer) # for summary table
library(janitor) # for tabyl
library(tigris) # for shapefiles
library(sf) # for spatial joins
library(leaflet) # for map
library(viridis)

nri <- read_csv("data/fema_nri_cville_tract.csv")
cvillefips <- c("540", "003", "065", "079", "109", "125")
```

## Data Source

Source: FEMA, National Risk Index, October 2020 release.

* Download URL: [https://nri-data-downloads.s3.amazonaws.com/NRI_Table_CensusTracts.zip](https://nri-data-downloads.s3.amazonaws.com/NRI_Table_CensusTracts.zip)
* About: [https://hazards.geoplatform.gov/portal/apps/MapSeries/index.html?appid=ddf915a24fb24dc8863eed96bc3345f8](https://hazards.geoplatform.gov/portal/apps/MapSeries/index.html?appid=ddf915a24fb24dc8863eed96bc3345f8)

## Variable descriptions
```{r}
glimpse(nri)
```

Observations are census tract estimates of...

* Population, building value, agricultural value within tract, area of tract
* Natural hazards include: CFLD - coastal flooding, DRGT - drought, HWAV - heat wave, HRCN - hurricane, RFLD - riverine flooding, SWND - strong wind
* Hazard measures include: EVNTS - number of events in recording period, AFREQ - annualized frequency (# events/# years in recording period)
* Exposure measures include: EXPB - building value exposure, EXPP - population exposure, EXPE - population equivalence exposure, EXPA - agricultural value exposure
* Historic loss ratio measures include: HLRB - historic loss ratio for building value, HLRA - historicla loss ratio for agriculture, HLRP - historical loss ratio for population, HLRR - historic loss ratio overall

## Summaries 

5-number summaries of (non-missing) numeric variables (remove tract identifiers)

```{r}
nri %>% select(-c(OID_:STATEFIPS, COUNTYTYPE:TRACTFIPS, NRI_VER)) %>% 
  select(where(~is.numeric(.x) && !is.na(.x))) %>% 
  as.data.frame() %>% 
  stargazer(., type = "text", title = "Summary Statistics", digits = 0,
            summary.stat = c("mean", "sd", "min", "median", "max"))
```

Summaries of (non-missing) character variables (remove tract identifiers)

```{r}
nri %>% select(-c(OID_:STATEFIPS, COUNTYTYPE:TRACTFIPS, NRI_VER)) %>% 
  select(where (~is.character(.x))) %>% map(tabyl)
```

##  Visual distribution
Via a grouped series of histograms

### Tract assets
```{r}
nri %>% select(TRACTFIPS:AREA) %>% 
  pivot_longer(-TRACTFIPS, names_to = "measure", values_to = "value") %>% 
  ggplot(aes(x = value, fill = measure)) + 
  scale_fill_viridis(option = "plasma", discrete = TRUE, guide = FALSE) +
  geom_histogram() + 
  facet_wrap(~measure, scales = "free")
```

### Tract hazards: Drought
```{r}
nri %>% select(contains("DRGT"), TRACTFIPS) %>% select(-contains("HLRR")) %>% 
  pivot_longer(-TRACTFIPS, names_to = "measure", values_to = "value") %>% 
  ggplot(aes(x = value, fill = measure)) + 
  geom_histogram() + 
  scale_fill_viridis(option = "plasma", discrete = TRUE, guide = FALSE) +
  facet_wrap(~measure, scales = "free")
```

### Tract hazards: Heat Wave
```{r}
nri %>% select(contains("HWAV"), TRACTFIPS) %>% select(-contains("HLRR")) %>% 
  pivot_longer(-TRACTFIPS, names_to = "measure", values_to = "value") %>% 
  ggplot(aes(x = value, fill = measure)) + 
  geom_histogram() + 
  scale_fill_viridis(option = "plasma", discrete = TRUE, guide = FALSE) +
  facet_wrap(~measure, scales = "free")
```

### Tract hazards: Hurricane
```{r}
nri %>% select(contains("HRCN"), TRACTFIPS) %>% select(-contains("HLRR")) %>% 
  pivot_longer(-TRACTFIPS, names_to = "measure", values_to = "value") %>% 
  ggplot(aes(x = value, fill = measure)) + 
  geom_histogram() + 
  scale_fill_viridis(option = "plasma", discrete = TRUE, guide = FALSE) +
  facet_wrap(~measure, scales = "free")
```

### Tract hazards: Riverine Flooding
```{r}
nri %>% select(contains("RFLD"), TRACTFIPS) %>% select(-contains("HLRR")) %>% 
  pivot_longer(-TRACTFIPS, names_to = "measure", values_to = "value") %>% 
  ggplot(aes(x = value, fill = measure)) + 
  geom_histogram() + 
  scale_fill_viridis(option = "plasma", discrete = TRUE, guide = FALSE) +
  facet_wrap(~measure, scales = "free")
```

### Tract hazards: Strong Wind
```{r}
nri %>% select(contains("SWND"), TRACTFIPS) %>% select(-contains("HLRR")) %>% 
  pivot_longer(-TRACTFIPS, names_to = "measure", values_to = "value") %>% 
  ggplot(aes(x = value, fill = measure)) + 
  geom_histogram() + 
  scale_fill_viridis(option = "plasma", discrete = TRUE, guide = FALSE) +
  facet_wrap(~measure, scales = "free")
```


## Maps

```{r, echo = FALSE}
cville_tracts <- readRDS("data/cville_tracts.RDS")

cville_nri <- cville_tracts %>% 
  left_join(nri, by = c("TRACTCE" = "TRACT"))

cville_nri <- st_transform(cville_nri, crs = 4326) # to WGS84, given error
```

### Droughts
```{r}
# DRGT
pal <- colorNumeric("plasma", reverse = TRUE, domain = cville_nri$DRGT_AFREQ) # viridis

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = cville_nri,
              fillColor = ~pal(DRGT_AFREQ),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("Tract Number: ", cville_nri$NAME, "<br>",
                             "Ann. Freq.: ", round(cville_nri$DRGT_AFREQ, 2))
  ) %>% 
  addLegend("bottomright", pal = pal, values = cville_nri$DRGT_AFREQ, 
            title = "Drought-#/year", opacity = 0.7)
```

### Heat Wave
```{r}
# HWAV
pal <- colorNumeric("plasma", reverse = TRUE, domain = cville_nri$HWAV_AFREQ) # viridis

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = cville_nri,
              fillColor = ~pal(HWAV_AFREQ),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("Tract Number: ", cville_nri$NAME, "<br>",
                             "Ann. Freq.: ", round(cville_nri$HWAV_AFREQ, 2))
  ) %>% 
  addLegend("bottomright", pal = pal, values = cville_nri$HWAV_AFREQ, 
            title = "Heat Wave-#/year", opacity = 0.7)
```

### Hurricane
```{r}
# HRCN
pal <- colorNumeric("plasma", reverse = TRUE, domain = cville_nri$HRCN_AFREQ) # viridis

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = cville_nri,
              fillColor = ~pal(HRCN_AFREQ),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("Tract Number: ", cville_nri$NAME, "<br>",
                             "Ann. Freq.: ", round(cville_nri$HRCN_AFREQ, 2))
  ) %>% 
  addLegend("bottomright", pal = pal, values = cville_nri$HRCN_AFREQ, 
            title = "Hurricane-#/year", opacity = 0.7)
```

### Riverine Flooding
```{r}
# RFLD
pal <- colorNumeric("plasma", reverse = TRUE, domain = cville_nri$RFLD_AFREQ) # viridis

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = cville_nri,
              fillColor = ~pal(RFLD_AFREQ),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("Tract Number: ", cville_nri$NAME, "<br>",
                             "Ann. Freq.: ", round(cville_nri$RFLD_AFREQ, 2))
  ) %>% 
  addLegend("bottomright", pal = pal, values = cville_nri$RFLD_AFREQ, 
            title = "Riverine Flooding-#/year", opacity = 0.7)
```


### Strong Wind
```{r}
# SWND
pal <- colorNumeric("plasma", reverse = TRUE, domain = cville_nri$SWND_AFREQ) # viridis

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = cville_nri,
              fillColor = ~pal(SWND_AFREQ),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("Tract Number: ", cville_nri$NAME, "<br>",
                             "Ann. Freq.: ", round(cville_nri$SWND_AFREQ, 2))
  ) %>% 
  addLegend("bottomright", pal = pal, values = cville_nri$SWND_AFREQ, 
            title = "Strong Wind-#/year", opacity = 0.7)
```

## Nota Bene

* Coastal flooding not relevant for this region.
* Most hazard overall ratings are very low or no ratings, except for drought; relative to larger national variation, these do not represent high risks.
* Several hazard rates are dominated by regional effects, with little variation within th region.