---
title: "FEMA National Risk Index"
subtitle: "Data for Charlottesville Region"
author: "Michele Claibourn"
date: "11/13/2021"
output: 
  html_document: 
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# 0. Load libraries and data
library(tidyverse)
library(stargazer) # for summary table
library(sf) # for spatial joins
library(leaflet) # for map
library(viridis)
library(googlesheets4)

nri <- read_csv("../data/fema_nri_cville_tract.csv")
# cvillefips <- c("540", "003", "065", "079", "109", "125")
meta <- read_sheet("https://docs.google.com/spreadsheets/d/1nqm3DuVXD1ObbVe_deacvT7uSLdBXfQJo3mkbqDwrVo/edit#gid=1573436636")
```

## Data Source

Source: FEMA, National Risk Index, October 2020 release.

* Download URL: [https://nri-data-downloads.s3.amazonaws.com/NRI_Table_CensusTracts.zip](https://nri-data-downloads.s3.amazonaws.com/NRI_Table_CensusTracts.zip)
* To learn more: [National Risk Index Primer, FEMA, December 2020](https://www.fema.gov/sites/default/files/documents/fema_national-risk-index_primer.pdf)
* To learn more: [National Risk Index Platform and Data](https://hazards.geoplatform.gov/portal/apps/MapSeries/index.html?appid=ddf915a24fb24dc8863eed96bc3345f8)
* Data is from the July 2021 release. There is no set update schedule. 


## About the data

The National Risk Index is intended to provide a view of the natural hazard risk within communities. While FEMA includes information on 18 natural hazards, we focus on six -- coastal flooding, drought, heat wave, hurricane, riverine flooding, and strong wind -- pulling measures on 

* frequency (measuring the number of events or event days during a reporting period and the estimated annualized frequency or probability), 
* exposure (measuring the building value, people, or agricultural value exposed to the natural hazard event), and 
* historic loss ratio (measuring the proportion of building value, people, or agricultural value that has been historically impacted by the natural hazard).

The NRI uses data on natural hazards from multiple sources and estimates natural hazard frequency, exposure, and historic loss at the census tract level.


## Variable descriptions
```{r}
glimpse(nri)
```

Observations are census tract estimates of...

* Population, building value, agricultural value, and area within tract
* Natural hazards include: CFLD - coastal flooding, DRGT - drought, HWAV - heat wave, HRCN - hurricane, RFLD - riverine flooding, SWND - strong wind
* Hazard measures include: EVNTS - number of events in recording period, AFREQ - annualized frequency (# events/# years in recording period)
* Exposure measures include: EXPB - building value exposure, EXPP - population exposure, EXPE - population equivalence exposure, EXPA - agricultural value exposure
* Historic loss ratio measures include: HLRB - historic loss ratio for building value, HLRA - historicla loss ratio for agriculture, HLRP - historical loss ratio for population, HLRR - historic loss ratio overall

## Summaries 

5-number summaries of (non-missing) numeric variables (remove tract identifiers)

```{r}
nri %>% select(-c(OID_:STATEFIPS, COUNTYTYPE:TRACTFIPS, NRI_VER)) %>% 
  select(where(~is.numeric(.x) && !is.na(.x))) %>% 
  as.data.frame() %>% 
  stargazer(., type = "text", title = "Summary Statistics", digits = 0,
            summary.stat = c("mean", "sd", "min", "median", "max"))
```


##  Visual distribution{.tabset}
Frequency distribution across tracts:

### Tract assets
```{r}
nri %>% select(TRACTFIPS:AREA) %>% 
  pivot_longer(-TRACTFIPS, names_to = "measure", values_to = "value") %>% 
  ggplot(aes(x = value, fill = measure)) + 
  scale_fill_viridis(option = "plasma", discrete = TRUE, guide = FALSE) +
  geom_histogram() + 
  facet_wrap(~measure, scales = "free")
```

```{r, results = "asis"}
meta %>% 
  filter(varname %in% c("POPULATION", "BUILDVALUE", "AGRIVALUE")) %>%
  mutate(label = paste0(varname, ": ", about)) %>% 
  select(label) %>% 
  as.list()
```

### Tract hazards: Drought
```{r}
vars <- nri %>% select(contains("DRGT"), -contains("HLRR")) %>% names()

nri %>% select(all_of(vars), TRACTFIPS) %>% 
  pivot_longer(-TRACTFIPS, names_to = "measure", values_to = "value") %>% 
  ggplot(aes(x = value, fill = measure)) + 
  geom_histogram() + 
  scale_fill_viridis(option = "plasma", discrete = TRUE, guide = FALSE) +
  facet_wrap(~measure, scales = "free")
```

```{r, results = "asis"}
meta %>%
  filter(varname %in% vars, !(str_detect(about, "REMOVE"))) %>% 
  mutate(label = paste0(varname, ": ", about)) %>% 
  select(label) %>% 
  as.list()
```

### Tract hazards: Heat Wave
```{r}
vars <- nri %>% select(contains("HWAV"), -contains("HLRR")) %>% names()

nri %>% select(all_of(vars), TRACTFIPS) %>% 
  pivot_longer(-TRACTFIPS, names_to = "measure", values_to = "value") %>% 
  ggplot(aes(x = value, fill = measure)) + 
  geom_histogram() + 
  scale_fill_viridis(option = "plasma", discrete = TRUE, guide = FALSE) +
  facet_wrap(~measure, scales = "free")
```

```{r, results = "asis"}
meta %>%
  filter(varname %in% vars, !(str_detect(about, "REMOVE"))) %>% 
  mutate(label = paste0(varname, ": ", about)) %>% 
  select(label) %>% 
  as.list()
```

### Tract hazards: Hurricane
```{r}
vars <- nri %>% select(contains("HRCN"), -contains("HLRR")) %>% names()

nri %>% select(all_of(vars), TRACTFIPS) %>% 
  pivot_longer(-TRACTFIPS, names_to = "measure", values_to = "value") %>% 
  ggplot(aes(x = value, fill = measure)) + 
  geom_histogram() + 
  scale_fill_viridis(option = "plasma", discrete = TRUE, guide = FALSE) +
  facet_wrap(~measure, scales = "free")
```

```{r, results = "asis"}
meta %>%
  filter(varname %in% vars, !(str_detect(about, "REMOVE"))) %>% 
  mutate(label = paste0(varname, ": ", about)) %>% 
  select(label) %>% 
  as.list()
```

### Tract hazards: Riverine Flooding
```{r}
vars <- nri %>% select(contains("RFLD"), -contains("HLRR")) %>% names()

nri %>% select(all_of(vars), TRACTFIPS) %>% 
  pivot_longer(-TRACTFIPS, names_to = "measure", values_to = "value") %>% 
  ggplot(aes(x = value, fill = measure)) + 
  geom_histogram() + 
  scale_fill_viridis(option = "plasma", discrete = TRUE, guide = FALSE) +
  facet_wrap(~measure, scales = "free")
```

```{r, results = "asis"}
meta %>%
  filter(varname %in% vars, !(str_detect(about, "REMOVE"))) %>% 
  mutate(label = paste0(varname, ": ", about)) %>% 
  select(label) %>% 
  as.list()
```

### Tract hazards: Strong Wind
```{r}
vars <- nri %>% select(contains("SWND"), -contains("HLRR")) %>% names()

nri %>% select(all_of(vars), TRACTFIPS) %>% 
  pivot_longer(-TRACTFIPS, names_to = "measure", values_to = "value") %>% 
  ggplot(aes(x = value, fill = measure)) + 
  geom_histogram() + 
  scale_fill_viridis(option = "plasma", discrete = TRUE, guide = FALSE) +
  facet_wrap(~measure, scales = "free")
```

```{r, results = "asis"}
meta %>%
  filter(varname %in% vars, !(str_detect(about, "REMOVE"))) %>% 
  mutate(label = paste0(varname, ": ", about)) %>% 
  select(label) %>% 
  as.list()
```


## Maps{.tabset}
Variation across tracts

```{r, echo = FALSE}
cville_tracts <- readRDS("../data/cville_tracts.RDS")

cville_nri <- cville_tracts %>% 
  left_join(nri, by = c("TRACTCE" = "TRACT"))

cville_nri <- st_transform(cville_nri, crs = 4326) # to WGS84, given error
```

### Droughts
```{r}
# DRGT
pal <- colorNumeric("plasma", reverse = TRUE, domain = cville_nri$DRGT_AFREQ) # viridis

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = cville_nri,
              fillColor = ~pal(DRGT_AFREQ),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("Tract Number: ", cville_nri$NAME, "<br>",
                             "Ann. Freq.: ", round(cville_nri$DRGT_AFREQ, 2))
  ) %>% 
  addLegend("bottomright", pal = pal, values = cville_nri$DRGT_AFREQ, 
            title = "Drought-#/year", opacity = 0.7)
```

```{r, results = "asis"}
meta %>%
  filter(varname == "DRGT_AFREQ") %>% 
  select(about) %>% 
  as.list()
```

### Heat Wave
```{r}
# HWAV
pal <- colorNumeric("plasma", reverse = TRUE, domain = cville_nri$HWAV_AFREQ) # viridis

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = cville_nri,
              fillColor = ~pal(HWAV_AFREQ),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("Tract Number: ", cville_nri$NAME, "<br>",
                             "Ann. Freq.: ", round(cville_nri$HWAV_AFREQ, 2))
  ) %>% 
  addLegend("bottomright", pal = pal, values = cville_nri$HWAV_AFREQ, 
            title = "Heat Wave-#/year", opacity = 0.7)
```

```{r, results = "asis"}
meta %>%
  filter(varname == "HWAV_AFREQ") %>% 
  select(about) %>% 
  as.list()
```

### Hurricane
```{r}
# HRCN
pal <- colorNumeric("plasma", reverse = TRUE, domain = cville_nri$HRCN_AFREQ) # viridis

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = cville_nri,
              fillColor = ~pal(HRCN_AFREQ),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("Tract Number: ", cville_nri$NAME, "<br>",
                             "Ann. Freq.: ", round(cville_nri$HRCN_AFREQ, 2))
  ) %>% 
  addLegend("bottomright", pal = pal, values = cville_nri$HRCN_AFREQ, 
            title = "Hurricane-#/year", opacity = 0.7)
```

```{r, results = "asis"}
meta %>%
  filter(varname == "HRCN_AFREQ") %>% 
  select(about) %>% 
  as.list()
```

### Riverine Flooding
```{r}
# RFLD
pal <- colorNumeric("plasma", reverse = TRUE, domain = cville_nri$RFLD_AFREQ) # viridis

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = cville_nri,
              fillColor = ~pal(RFLD_AFREQ),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("Tract Number: ", cville_nri$NAME, "<br>",
                             "Ann. Freq.: ", round(cville_nri$RFLD_AFREQ, 2))
  ) %>% 
  addLegend("bottomright", pal = pal, values = cville_nri$RFLD_AFREQ, 
            title = "Riverine Flooding-#/year", opacity = 0.7)
```

```{r, results = "asis"}
meta %>%
  filter(varname == "RFLD_AFREQ") %>% 
  select(about) %>% 
  as.list()
```


### Strong Wind
```{r}
# SWND
pal <- colorNumeric("plasma", reverse = TRUE, domain = cville_nri$SWND_AFREQ) # viridis

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = cville_nri,
              fillColor = ~pal(SWND_AFREQ),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("Tract Number: ", cville_nri$NAME, "<br>",
                             "Ann. Freq.: ", round(cville_nri$SWND_AFREQ, 2))
  ) %>% 
  addLegend("bottomright", pal = pal, values = cville_nri$SWND_AFREQ, 
            title = "Strong Wind-#/year", opacity = 0.7)
```

```{r, results = "asis"}
meta %>%
  filter(varname == "SWND_AFREQ") %>% 
  select(about) %>% 
  as.list()
```

## Nota Bene

* Coastal flooding not relevant for this region (included for consistency with Eastern Shore work).
* Most hazard overall ratings are very low or no ratings, except for drought; relative to larger national variation, these do not represent high risks.
* Several hazard rates are dominated by regional effects, with little variation within the region.