---
title: "Charlottesville Region HMDA"
author: "Lee LeBoeuf"
date: "10/15/2021"
output: 
  html_document: 
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

# invisible(lapply(list('rgdal', 'lodes', 'psych', 'reshape2', 'transformer', 'gifski', 'dygraphs', 'xts', 'sp', 'geosphere', 'stringr', 'ggrepel', 'gganimate')))


invisible(lapply(list('tidyverse', 'stargazer', 'janitor', 'tigris', 'sf', 'leaflet', 'rcartocolor', 'RColorBrewer', 'viridis', 'googlesheets4', 'gganimate'),
                 function(pkg) library(pkg, character.only = TRUE)))

meta <- read_sheet("https://docs.google.com/spreadsheets/d/1nqm3DuVXD1ObbVe_deacvT7uSLdBXfQJo3mkbqDwrVo/edit?usp=sharing", sheet = "HMDA", gs4_deauth())

# main data
cvldat <- read.csv("../data/hmda_cville_tract.csv")
cvldat <- cvldat %>% 
  mutate(census_tract = as.character(census_tract),
         countyfips = str_sub(census_tract, 3,5))


# spatial data
shape <- readRDS('../data/cville_tracts.RDS')
shape <- st_transform(shape, crs = 4326) # to WGS84, given error
```

## Data Source
* These data come from the Consumer Financial Protection Bureau and Federal Financial Institutions Examinations Council. All data were collected as part of the Home Mortgage Disclosure Act (HMDA). 
* New data is released biannually on a year and half delay from when it was collected. New data can be downloaded from the Federal Financial Institutions Examinations Council [here](https://ffiec.cfpb.gov/data-browser/data/2018?category=states&items=VA). 
* Data from 2017 and earlier can be downloaded from the Consumer Financial Protection Bureau [here](https://www.consumerfinance.gov/data-research/hmda/historic-data/?geo=va&records=all-records&field_descriptions=labels)
* Data presented here are from 2007-2020 (mostly tract averages across those years). 

## Data specifics 
* The HMDA requires financial institutions to maintain, report, and publicly disclose information about mortgages. 
* New data is released annually, typically in the summer months, on a year delay (2020 data was released in June of 2021).
* Areas included in these data: Charlottesville City (51540), Albemarle County (51003), Fluvanna (51065), Greene (51079), Louisa (51109), Nelson (51125)
* The data show below include __only__ home purchase loans, meaning that mortgages for home improvement or refinancing are excluded. 
* An important limitation to note is that the FFEIC allows banks to submit data with an error threshold as high as 10% for community banks and credit unions. The loan amounts and applicant income values appear prone to entry error at the individual level--a high proportion of values seemed too high to be realistic. Therefore, the tract level summaries of these variables (median loan amounts and median applicant income in the tract) should be interpreted with extreme caution. 
    
## Variable descriptions and Summaries
```{r}
# meta %>% 
#   filter(su_tract == 1) %>%
#   select(varname, about) %>% as.list()

glimpse(cvldat)

cvldat[which(cvldat$year == 2020),] %>% select(total_apps, overall_denial_rate, white_denial_rate, black_denial_rate, hislat_denial_rate, perc_conventional, perc_govern_backed) %>% 
  select(where(~is.numeric(.x))) %>% 
  as.data.frame() %>% 
  stargazer(., type = "text", title = "Summary Statistics", digits = 2,
            summary.stat = c("mean", "sd", "min", "median", "max"))
```

## Visual distribution 
* The following distributions include data only from the year 2020. 
```{r}
longdat <- cvldat[which(cvldat$year == 2020),] %>% select(c(census_tract, total_apps, overall_denial_rate, white_denial_rate, black_denial_rate, hislat_denial_rate, perc_conventional, perc_govern_backed)) %>% pivot_longer(-census_tract, names_to = "measure", values_to = "value") 

longdat$measure <- factor(longdat$measure,
                         levels = c("total_apps", "overall_denial_rate", "white_denial_rate", "black_denial_rate", "hislat_denial_rate", "perc_conventional", "perc_govern_backed"))

longdat %>%
  ggplot(aes(x = value, fill = measure)) +
  scale_fill_viridis(option = "plasma", discrete = TRUE, guide = FALSE) +
  geom_histogram() + 
  facet_wrap(~measure, scales = "free")
```

```{r, results = "asis"}
meta %>% 
  filter(varname %in% c("total_apps", "overall_denial_rate", "white_denial_rate", "black_denial_rate", "hislat_denial_rate", "perc_conventional", "perc_govern_backed")) %>%
  mutate(label = paste0(varname, ": ", about)) %>% 
  select(label) %>% 
  as.list()
```

## Distribution over Time{.tabset}

### Total number of applications
```{r}
cvldat %>% 
  group_by(countyfips, year) %>% 
  summarize(total_apps = sum(total_apps)) %>% 
  ggplot(aes(x = year, y = total_apps, color = countyfips)) +
  geom_line()
```

### Application Denials
```{r}
cvldat %>% 
  group_by(countyfips, year) %>% 
  mutate(denial_rate = mean(na.omit(overall_denial_rate))) %>%
  summarize(denial_rate = mean(na.omit(overall_denial_rate))) %>% 
  ggplot(aes(x = year, y = denial_rate, color = countyfips)) +
  geom_line()
```
   
## Mapping the data 
```{r, echo = FALSE}
mapdat <- left_join(shape, cvldat, by = c('GEOID'= "census_tract"))
```

### Total number of applications in 2020{.tabset}

#### Total Applications
```{r }
mapdat2020 <- mapdat[which(mapdat$year == 2020),]

pal <- colorNumeric("plasma", reverse = TRUE, domain = mapdat2020$total_apps)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat2020,
              fillColor = ~pal(mapdat2020$total_apps),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("GEOID: ", mapdat2020$GEOID, "<br>",
                             "Number of applications: ", mapdat2020$total_apps)
  ) %>% 
  addLegend("bottomright", pal = pal, values = (mapdat2020$total_apps), 
            title = "Total number of <br>mortgage applications <br> in 2020", opacity = 0.7)
```

#### Applications by Race: White Applicants
```{r }
pal <- colorNumeric("plasma", reverse = TRUE, domain = mapdat2020$appRace_White)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat2020,
              fillColor = ~pal(mapdat2020$appRace_White),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("GEOID: ", mapdat2020$GEOID, "<br>",
                             "Number of applications: ", mapdat2020$appRace_White)
  ) %>% 
  addLegend("bottomright", pal = pal, values = (mapdat2020$appRace_White), 
            title = "Total 2020 mortgage<br>applications: White", opacity = 0.7)
```

#### Applications by Race: Black Applicants
```{r }
pal <- colorNumeric("plasma", reverse = TRUE, domain = mapdat2020$appRace_Black)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat2020,
              fillColor = ~pal(mapdat2020$appRace_Black),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("GEOID: ", mapdat2020$GEOID, "<br>",
                             "Number of applications: ", mapdat2020$appRace_Black)
  ) %>% 
  addLegend("bottomright", pal = pal, values = (mapdat2020$appRace_Black), 
            title = "Total 2020 mortgage<br>applications: Black", opacity = 0.7)
```

#### Applications by Race: Asian Applicants
```{r }
pal <- colorNumeric("plasma", reverse = TRUE, domain = mapdat2020$appRace_Asian)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat2020,
              fillColor = ~pal(mapdat2020$appRace_Asian),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("GEOID: ", mapdat2020$GEOID, "<br>",
                             "Number of applications: ", mapdat2020$appRace_Asian)
  ) %>% 
  addLegend("bottomright", pal = pal, values = (mapdat2020$appRace_Asian), 
            title = "Total 2020 mortgage<br>applications: Asian", opacity = 0.7)
```

#### Applications by Ethnicity: Hispanic Applicants
```{r }
pal <- colorNumeric("plasma", reverse = TRUE, domain = mapdat2020$appEth_HispLat)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat2020,
              fillColor = ~pal(mapdat2020$appEth_HispLat),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("GEOID: ", mapdat2020$GEOID, "<br>",
                             "Number of applications: ", mapdat2020$appEth_HispLat)
  ) %>% 
  addLegend("bottomright", pal = pal, values = (mapdat2020$appEth_HispLat), 
            title = "Total 2020 mortgage<br>applications: Hispanic", opacity = 0.7)
```

### Average mortage application denial rates from 2007-2020 {.tabset}

Each of the following maps show the average denial rates for the years 2007-2020

#### Overall denial rate {.tabset}
```{r }
mapdat <- mapdat %>%
  group_by(GEOID) %>%
  mutate(avg_overall_denial_rate = mean(na.omit(overall_denial_rate))) %>%
  ungroup()

pal <- colorNumeric("plasma", reverse = TRUE, domain = mapdat$avg_overall_denial_rate)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat,
              fillColor = ~pal(mapdat$avg_overall_denial_rate),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("GEOID: ", mapdat$GEOID, "<br>",
                             "Average overall app denial rate from 2007-2020: ", round(mapdat$avg_overall_denial_rate, 2))
  ) %>% 
  addLegend("bottomright", pal = pal, values = (mapdat$avg_overall_denial_rate), 
            title = "Average overall <br>app denial rate <br>from 2007-2020", opacity = 0.7)
```

#### Denial Rate by Race: White
```{r }
mapdat <- mapdat %>%
  group_by(GEOID) %>%
  mutate(avg_white_denial_rate = mean(na.omit(white_denial_rate))) %>%
  ungroup()

pal <- colorNumeric("plasma", reverse = TRUE, domain = mapdat$avg_white_denial_rate)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat,
              fillColor = ~pal(mapdat$avg_white_denial_rate),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("GEOID: ", mapdat$GEOID, "<br>",
                             "Average White app denial rate from 2007-2020: ", round(mapdat$avg_white_denial_rate, 2))
  ) %>% 
  addLegend("bottomright", pal = pal, values = (mapdat$avg_white_denial_rate), 
            title = "Average White <br>app denial rate <br>from 2007-2020", opacity = 0.7)
```

#### Denial Rate by Race: Black
```{r }
mapdat <- mapdat %>%
  group_by(GEOID) %>%
  mutate(avg_black_denial_rate = mean(na.omit(black_denial_rate))) %>%
  ungroup()

pal <- colorNumeric("plasma", reverse = TRUE, domain = (mapdat$avg_black_denial_rate))

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat,
              fillColor = ~pal((mapdat$avg_black_denial_rate)),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("GEOID: ", mapdat$GEOID, "<br>",
                             "Average Black app denial rate from 2007-2020: ", round(mapdat$avg_black_denial_rate, 2))
  ) %>% 
  addLegend("bottomright", pal = pal, values = (mapdat$avg_black_denial_rate), 
            title = "Average Black <br>app denial rate <br>from 2007-2020", opacity = 0.7)
```

#### Denial Rate by Ethnicity: Hispanic
```{r }
mapdat <- mapdat %>%
  group_by(GEOID) %>%
  mutate(avg_hislat_denial_rate = mean(na.omit(hislat_denial_rate))) %>%
  ungroup()

pal <- colorNumeric("plasma", reverse = TRUE, domain = (mapdat$avg_hislat_denial_rate))

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat,
              fillColor = ~pal((mapdat$avg_hislat_denial_rate)),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("GEOID: ", mapdat$GEOID, "<br>",
                             "Average His/Lat app denial rate from 2007-2020: ", round(mapdat$avg_hislat_denial_rate, 2))
  ) %>% 
  addLegend("bottomright", pal = pal, values = (mapdat$avg_hislat_denial_rate), 
            title = "Average Hispanic/Latino <br>app denial rate <br>from 2007-2020", opacity = 0.7)
```

### Government backed mortgages {.tabset}

Government back mortages are mortgages insured by the USDA, VA, or FHA

#### Percent of government backed loans from 2007-2020
```{r }
mapdat <- mapdat %>%
  group_by(GEOID) %>%
  mutate(avg_percgovbacked = mean(na.omit(perc_govern_backed))) %>%
  ungroup()

pal <- colorNumeric("plasma", reverse = TRUE, domain = mapdat$avg_percgovbacked)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat,
              fillColor = ~pal(mapdat$avg_percgovbacked),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("GEOID: ", mapdat$GEOID, "<br>",
                             "Average % of gov-backed mortgages <br> from 2006-2020: ", round(mapdat$avg_percgovbacked, 2))
  ) %>% 
  addLegend("bottomright", pal = pal, values = (mapdat$avg_percgovbacked), 
            title = "Average % of <br>gov-backed mortgages <br>from 2007-2020", opacity = 0.7)
```

#### Percent of government backed loans in 2020 
```{r }
pal <- colorNumeric("plasma", reverse = TRUE, domain = mapdat2020$perc_govern_backed)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = mapdat2020,
              fillColor = ~pal(mapdat2020$perc_govern_backed),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.6,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T
              ),
              popup = paste0("GEOID: ", mapdat2020$GEOID, "<br>",
                             "Percent of gov-backed <br>mortages in 2020: ", round(mapdat2020$perc_govern_backed, 2))
              ) %>% 
  addLegend("bottomright", pal = pal, values = (mapdat2020$perc_govern_backed), 
            title = "Percent of <br>gov-backed <br>mortages in 2020", opacity = 0.7)
```


```{r}
# animatemapdat <- st_as_sf(mapdat)
# 
# animatemapdat$year = as.numeric(animatemapdat$year)
# 
# animatemapdat <-  animatemapdat %>% filter_at(vars(NAME, geometry, perc_govern_backed),all_vars(!is.na(.)))

# cville1 <- 
#   ggplot(animatemapdat) +
#   geom_sf(aes(fill = perc_govern_backed), color = "black", alpha = .9, na.rm = TRUE) +
#   scale_fill_fermenter(palette = "Blues", direction = 1,   type = "seq", n.breaks = 7) +
#   theme_void() +
#   guides(fill = guide_colourbar(title.position="top", title.hjust = 0.5, barwidth = 1))  + 
#   labs(fill = "Percent of mortgages backed by gov.", title = 'Year:{frame_time}',
#        caption = "Percent of approved mortgage apps backed by the USDA, VA, or FHA") + 
#   transition_time(as.integer(year)) +
#   ease_aes('linear') 

# animate(cville1, fps = 1, nframes = 13)
```

