---
title: "NOAA Temperature Data"
subtitle: "Data for Charlottesville Region"
author: "Tolu Odukoya"
date: "7/4/2021"
output: 
  html_document: 
    toc: true
    toc_float: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
# 0. Load libraries and data
library(tidyverse)
library(stargazer) # for summary table
library(janitor) # for tabyl
library(tigris) # for shapefiles
library(sf) # for spatial joins
library(leaflet) # for map
library(viridis)
library(googlesheets4)
library(RColorBrewer)
library(ggplot2)
library(gganimate)
library(gapminder)
library(transformr)
library(gifski)
library(ggrepel)
library(hrbrthemes)
library(devtools)
library(climatestripes)
library(lubridate)
noaa <- read_csv("noaa_cville_county.csv")
cvillefips <- c("540", "003", "065", "079", "109", "125")
meta <- read_sheet("https://docs.google.com/spreadsheets/d/1nqm3DuVXD1ObbVe_deacvT7uSLdBXfQJo3mkbqDwrVo/edit#gid=1573436636", sheet = "noaa")
```

```{r load and clean new data shp, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

cville_sf <- read_sf("/Users/msbugatti/Documents/Intership 2021/NOAA Temp Work/counties")
cville_sf <- cville_sf %>% filter(FIPS %in% c(51540, 51003, 51065, 51079, 51109, 51125))
cville_sf <- cville_sf %>%
                      dplyr::select(NAME, STATE_FIPS, CNTY_FIPS, FIPS, geometry, AREA)
cville_sf1 <- left_join(cville_sf, noaa, by ="CNTY_FIPS")
```


```{r transpose the data, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

cville_sf1 <- st_as_sf(cville_sf1)
cville_sf1 <- cbind(cville_sf1, st_coordinates(st_centroid(cville_sf1)))
```

```{r remove na, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

cville_sf1 <- cville_sf1 %>%
                      dplyr::select(NAME, Year, CNTY_FIPS, X, Y,
                                geometry, everything())
cville_sf1$Year = as.numeric(cville_sf1$Year)
cville_sf1 <-  cville_sf1 %>% filter_at(vars(Jul:Dec),all_vars(!is.na(.)))
```

## Data Source

Source: NOAA's Climate Divisional Database (nClimDiv), June 2021 release

* Download URL: [https://www.ncei.noaa.gov/pub/data/cirs/climdiv/counties.zip](https://www.ncei.noaa.gov/pub/data/cirs/climdiv/counties.zip)
* Download URL: [https://www.ncei.noaa.gov/pub/data/cirs/climdiv/climdiv-tmaxcy-v1.0.0-20210604](https://www.ncei.noaa.gov/pub/data/cirs/climdiv/climdiv-tmaxcy-v1.0.0-20210604)

### The Data Used Here

The Climate Divisional Dataset from the National Center For Environmental Information began as the only long-term temporally and spatially complete dataset from which to generate historical climate analyses (1895-2013) for the contiguous United States (CONUS). It was originally developed for climate division, statewide, regional, national, and population-weighted monitoring of drought, temperature, precipitation, and heating/cooling degree day values. 

There are 344 climate divisions in the CONUS. For each climate division, monthly station temperature and precipitation values are computed from the daily observations. The divisional values are weighted by area to compute statewide values and the statewide values are weighted by area to compute regional values. (Karl and Koss, 1984).

The data here is the county data from all the states in the contigous United States. While NOAA includes information on a variety of climate measures, we are focusing on the maximum temperature of the counties of interest. 

To Learn More, See: 

* [ National Center For Environmental Information and Data ](https://www.ncdc.noaa.gov/monitoring-references/maps/us-climate-divisions.php#grdd)


## Variable Descriptions

```{r}
glimpse(cville_sf1)
```

Observations are county level estimates of...

* Monthly maximum temperature of each county for years 1895-2021
* Yearly average maximum temperature for each county for those years. 

## Summaries 

5-number summaries of (non-missing) numeric variables (remove non-numeric observations)

```{r}
noaa %>% select(-c(CNTY_FIPS, Fipsyear, Year)) %>% 
  select(where(~is.numeric(.x) && !is.na(.x))) %>% 
  as.data.frame() %>% 
  stargazer(., type = "text", title = "Summary Statistics", digits = 0,
            summary.stat = c("mean", "sd", "min", "median", "max"))
```


## Maps

#### Maximum Temperature in July across Counties for all Years 

```{r}
cville_sf1a <- 
  ggplot(cville_sf1) +
  geom_sf(aes(fill = Jul), color = "black", alpha = .9, na.rm = TRUE) +
   geom_text_repel(data = cville_sf1, aes(X, Y, label = NAME), size = 4, nudge_x = 1, nudge_y = 0, fontface = "bold", hjust = 0.9) +
  scale_fill_fermenter(palette = "YlOrRd", direction = 1,   type = "seq", n.breaks = 7) +
     theme_void() +
  guides(fill =
           guide_colourbar(title.position="top", title.hjust = 0.5,
                           barwidth = 1)
  )  + 
  labs(fill = "Temperature ", title = 'Year: {frame_time}',
       caption = "Maximum Temperature in July for Charlottesville Counties") + 
 transition_time(as.integer(Year)) +
ease_aes('linear') 
```

```{r}
animate(cville_sf1a, fps = 1, detail = 1, nframes = 127)
```

```{r, results = "asis"}
meta %>%
  filter(varname == "Jul") %>% 
  select(about) %>% 
  as.list()
```

#### Warming Stripes Showing Maximum Temperature in July across Counties for all Years 
```{r warming1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

#select only the annual temperature and year column
cville_sf1_jl <- select(cville_sf1, NAME, Year, Jul)

#rename the temperature column
cville_sf1_jl <- rename(cville_sf1_jl, ta = Jul)

```

```{r warming2, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

cville_sf1_jl <- mutate(cville_sf1_jl, date = str_c(Year, "01-01", sep = "-") %>% ymd())

```

```{r  warming3, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

cville_sf1_jlg <- filter(cville_sf1_jl, NAME == "Greene")
cville_sf1_jla <- filter(cville_sf1_jl, NAME == "Albemarle")
cville_sf1_jlc <- filter(cville_sf1_jl, NAME == "Charlottesville")
cville_sf1_jlf <- filter(cville_sf1_jl, NAME == "Fluvanna")
cville_sf1_jll <- filter(cville_sf1_jl, NAME == "Louisa")
cville_sf1_jln <- filter(cville_sf1_jl, NAME == "Nelson")
```

```{r  warming4, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

theme_strip <- theme_minimal()+
                 theme(axis.text.y = element_blank(),
                       axis.line.y = element_blank(),
                       axis.title = element_blank(),
                       panel.grid.major = element_blank(),
                       legend.title = element_blank(),
                       axis.text.x = element_text(vjust = 3),
                       panel.grid.minor = element_blank(),
                        plot.title = element_text(size = 14, face = "bold")
                       )


col_strip <- brewer.pal(11, "RdBu")
```

##### Greene County
```{r stripes1,fig.width=10, fig.height=3, message=FALSE, warning=FALSE}
 ggplot(cville_sf1_jlg,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Greene County July Maximum Temperature 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```

##### Louisa County
```{r stripes2,fig.width=10, fig.height=3, message=FALSE, warning=FALSE}
 ggplot(cville_sf1_jll,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Louisa County July Maximum Temperature 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```

##### Nelson County
```{r stripes3,fig.width=10, fig.height=3, message=FALSE, warning=FALSE}
 ggplot(cville_sf1_jln,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Nelson County July Maximum Temperature 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```

##### Albemarle County
```{r stripes4,fig.width=10, fig.height=3, message=FALSE, warning=FALSE}
 ggplot(cville_sf1_jla,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Albemarle County July Maximum Temperature 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```

##### Charlottesville County
```{r stripes5,fig.width=10, fig.height=3, message=FALSE, warning=FALSE}
 ggplot(cville_sf1_jlc,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Charlottesville County July Maximum Temperature 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```

##### Fluvanna County
```{r stripes6,fig.width=10, fig.height=3, message=FALSE, warning=FALSE}
 ggplot(cville_sf1_jlf,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Fluvanna County July Maximum Temperature 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```




#### Average Yearly Temperature across Counties

```{r}
cville_sf1b <- 
  ggplot(cville_sf1) +
  geom_sf(aes(fill = Avg_Temp), color = "black", alpha = .9, na.rm = TRUE) +
   geom_text_repel(data = cville_sf1, aes(X, Y, label = NAME), size = 4, nudge_x = 1, nudge_y = 0, fontface = "bold", hjust = 0.9) +
  scale_fill_fermenter(palette = "BuPu", direction = 1,   type = "seq", n.breaks = 6) +
     theme_void() +
  guides(fill =
           guide_colourbar(title.position="top", title.hjust = 0.5,
                           barwidth = 1)
  )  + 
  labs(fill = "Temperature ", title = 'Year: {frame_time}',
       caption = "Average Yearly Maximum 
       Temperature For Charlottesville Counties") + 
 transition_time(as.integer(Year)) +
ease_aes('linear') 
```

```{r}
animate(cville_sf1b, fps = 1, detail = 1, nframes = 127)
```


```{r, results = "asis"}
meta %>%
  filter(varname == "Avg_Temp") %>% 
  select(about) %>% 
  as.list()
```

#### Warming Stripes Showing Average Yearly Temperature across Counties
```{r  warming6, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

#select only the annual temperature and year column
cville_sf1_yr <- select(cville_sf1, NAME, Year, Avg_Temp)

#rename the temperature column
cville_sf1_yr <- rename(cville_sf1_yr, ta = Avg_Temp)

```

```{r  warming7, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

cville_sf1_yr <- mutate(cville_sf1_yr, date = str_c(Year, "01-01", sep = "-") %>% ymd())

```

```{r warming8, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

cville_sf1_yrg <- filter(cville_sf1_yr, NAME == "Greene")
cville_sf1_yra <- filter(cville_sf1_yr, NAME == "Albemarle")
cville_sf1_yrc <- filter(cville_sf1_yr, NAME == "Charlottesville")
cville_sf1_yrf <- filter(cville_sf1_yr, NAME == "Fluvanna")
cville_sf1_yrl <- filter(cville_sf1_yr, NAME == "Louisa")
cville_sf1_yrn <- filter(cville_sf1_yr, NAME == "Nelson")
```

##### Greene County
```{r stripes1a,fig.width=10, fig.height=3,  message=FALSE, warning=FALSE}
 ggplot(cville_sf1_yrg,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Greene County Yearly Maximum Temperature 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```

##### Louisa County
```{r stripes2a,fig.width=10, fig.height=3, message=FALSE, warning=FALSE}
 ggplot(cville_sf1_yrl,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Louisa County Yearly Maximum Temperature 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```

##### Nelson County
```{r stripes3a,fig.width=10, fig.height=3, message=FALSE, warning=FALSE}
 ggplot(cville_sf1_yrn,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Nelson County Yearly Maximum Temperature 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```

##### Albemarle County
```{r stripes4a,fig.width=10, fig.height=3,  message=FALSE, warning=FALSE}
 ggplot(cville_sf1_yra,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Albemarle County Yearly Maximum Temperature 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```

##### Charlottesville County
```{r stripes5a,fig.width=10, fig.height=3, message=FALSE, warning=FALSE}
 ggplot(cville_sf1_yrc,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Charlottesville County Yearly Maximum Temperature 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```

##### Fluvanna County
```{r stripes6a,fig.width=10, fig.height=3, message=FALSE, warning=FALSE}
 ggplot(cville_sf1_yrf,
             aes(x = date, y = 1, fill = ta))+
        geom_tile()+
           scale_x_date(date_breaks = "6 years",
                     date_labels = "%Y",
                     expand = c(0, 0))+
           scale_y_continuous(expand = c(0, 0))+
           scale_fill_gradientn(colors = rev(col_strip))+
             guides(fill = guide_colorbar(barwidth = 1))+
            labs(title = "Fluvanna County Yearly Maximum Temperature 1895-2020",
                caption = "Data: NOAA Surface Temperature Analysis")+
              theme_strip
```
