---
title: "Bivariate Choropleth"
author: "Michele"
date: "12/7/2021"
output: 
  html_document: 
    code_folding: hide
---

```{r start, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```


```{r setup}
library(tidyverse)
library(sf)
library(leaflet)
library(leafem)
library(biscale)
library(patchwork)
library(scales)

places <- read_csv("../data/cdcplaces_cville_tract.csv")
places <- places %>% 
  mutate(GEOID = as.character(locationname))

airquality <- read_csv("../data/airquality_cville_tract.csv")
airquality <- airquality %>% 
  mutate(GEOID = as.character(gid))

tracts <- readRDS("../data/cville_tracts.RDS")

df <- left_join(places, airquality, by = "GEOID")
df <- left_join(tracts, df)
```


## Using `geom_sf` and `biscale`
More on biscale: https://slu-opengis.github.io/biscale/articles/biscale.html

```{r biscale}
# create classes
df_bs <- bi_class(df, x = pm2_5_2016, y = Current_Asthma2018, style = "quantile", dim = 3)

map <- ggplot() +
  geom_sf(data = df_bs, 
          mapping = aes(fill = bi_class), color = "white", size = 0.1, show.legend = FALSE) +
  bi_scale_fill(pal = "DkBlue", dim = 3) +
  labs(title = "Air Quality and Asthma") +
  bi_theme()

# legend
legend <- bi_legend(pal = "DkBlue",
                    dim = 3,
                    xlab = "PM ",
                    ylab = "Asthma  ",
                    size = 8)

map + inset_element(legend, left = 0.7, bottom = 0.7, right = 1, top = 1)
```

## Use `geom_sf` and manually defined palette
Palette defined here: https://www.joshuastevens.net/cartography/make-a-bivariate-choropleth-map/ 

```{r palette}
# to use
bipal <- c("#e8e8e8", "#dfd0d6", "#be64ac", # A-1, A-2, A-3,
           "#ace4e4", "#a5add3", "#8c62aa", # B-1, B-2, B-3
           "#5ac8c8", "#5698b9", "#3b4994") # C-1, C-2, C-2

# to show
bipal2 <- c("#be64ac", "#8c62aa", "#3b4994",
            "#dfd0d6", "#a5add3", "#5698b9",
            "#e8e8e8", "#ace4e4", "#5ac8c8") 
            # A-3, B-3, C-3
            # A-2, B-2, C-2
            # A-1, B-1, C-1

show_col(bipal2)
```

Estimate terciles of variables and use above palette to map; recreate legend manually.
```{r manual}
# create class
df <- df %>%
  mutate(asthma = ntile(Current_Asthma2018, 3),
         pm25_2016 = ntile(pm2_5_2016, 3)) %>%
  mutate(pm25_2016 = if_else(pm25_2016 == 1, 'A', 
                          if_else(pm25_2016 == 2, 'B', 'C')),
         biclass = paste0(pm25_2016, asthma)) 

map2 <- ggplot(df) + 
  geom_sf(aes(fill = biclass), color = "white", size = 0.1, show.legend = FALSE) + 
  scale_fill_manual(values = bipal) +
  labs(title = "Air Quality and Asthma") +
  theme_void()

# make legend
bipal3 <- tibble(
  "3-3" = "#3b4994", # high pm2.5, high asthma
  "2-3" = "#8c62aa",
  "1-3" = "#be64ac", # low pm2.5, high asthma
  "3-2" = "#5698b9",
  "2-2" = "#a5add3", # medium pm2.5, medium asthma
  "1-2" = "#dfd0d6",
  "3-1" = "#5ac8c8", # high pm2.5, low asthma
  "2-1" = "#ace4e4",
  "1-1" = "#e8e8e8" # low pm2.5, low asthma
) %>%
  gather("group", "fill")

bipal3 <- bipal3 %>% 
  separate(group, into = c("pm2.5", "asthma"), sep = "-") %>%
  mutate(pm2.5 = as.integer(pm2.5),
         asthma = as.integer(asthma))

legend2 <- ggplot() +
  geom_tile(
    data = bipal3,
    mapping = aes(
      x = pm2.5,
      y = asthma,
      fill = fill)
  ) +
  scale_fill_identity() +
  labs(x = expression("Higher pm2.5" %->%  ""),
       y = expression("Higher asthma" %->% "")) +
  theme_void() +
  # make font small enough
  theme(
    axis.title = element_text(size = 6),
    axis.title.y = element_text(angle = 90)
  ) +
  # quadratic tiles
  coord_fixed()

# plot
map2 + inset_element(legend2, left = 0.8, bottom = 0.8, right = 1, top = 1)
```

## Use `leaflet`

Can't find a way to adapt the leaflet legend so saving the ggplot legend figure as an image and adding it in with `addLogo`.

```{r leaflet}
# save image
ggsave(plot = legend2, filename = "bivariate_legend.svg",
       width = 1, height = 1)

df_4326 <- st_transform(df, 4326)

factpal <- colorFactor(bipal, domain = df_4326$biclass)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = df_4326,
              fillColor = ~factpal(biclass),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.8,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T),
              popup = paste0("GEOID: ", df_4326$GEOID, "<br>",
                             "% with Asthma: ", 
                             df_4326$Current_Asthma2018, "<br>",
                             "PM_2.5 Levels: ", 
                             round(df_4326$pm2_5_2016,1))) %>% 
    addLogo("bivariate_legend.svg", src = "local",
          position = "topright", width = 100, height = 100,
          alpha = 0.8)
```

ARRRRRGH! In my R script, this worked, but I can't seem to get the legend image to appear in the R markdown version here! If I save the image online (outside of the script) and call it, does it work?

```{r leaflet2}
img <- "https://raw.githubusercontent.com/virginiaequitycenter/summer-sandbox/main/cville_region_collection/images/bivariate_legend.svg"

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = df_4326,
              fillColor = ~factpal(biclass),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.8,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T),
              popup = paste0("GEOID: ", df_4326$GEOID, "<br>",
                             "% with Asthma: ", 
                             df_4326$Current_Asthma2018, "<br>",
                             "PM_2.5 Levels: ", 
                             round(df_4326$pm2_5_2016,1))) %>% 
    addLogo(img,
          position = "topright", width = 100, height = 100,
          alpha = 0.8)
```

Yay? As that's totally useless to me for use in a more generalized way in an app.



