---
title: "Rendering locally saved legend in Leaflet map"
author: "jacob-gg"
date: "12/14/2021"
output:
  html_document:
      code_folding: hide
---

I’ve been looking into the cause of the issue Michele described in bichoropleth test doc (`bichoropleth.html`), where the custom Leaflet legend wouldn’t appear once the map was rendered (unless the legend was hosted on the web and called in). 

When I looked at the underlying HTML of the rendered map, the path to the legend appeared wrong to me:
`logo = addElement();logo.html('<img src=\"../graphs/bivariate_legend.svg\", width=100, height=100,`
`style=\"opacity:0.8;filter:alpha(opacity=80);\", ><\/a>');\n`

`../graphs/...` is a temporary directory created by the `leafem::addLogo()` function. I think the `../` relative path and/or the temporary graphs folder cause rendering issues. I was able to manually fix the problem by editing the HTML in a browser: When I switched the HTML above to a full local path (below), the legend appeared correctly.
`<img src="/Users/jacob_local_path/summer-sandbox/cville_region_collection/images/bivariate_legend.svg" ,=""` `style="opacity:0.8;filter:alpha(opacity=80);" width="100," height="100,">,`

To try and create a fix (at least a temporary one), I modified the `addLogo()` source code so that it no longer puts the custom legend in a temporary folder. The approach I've found that currently gets the legend to successfully render without needing to host it somewhere on the web is:

- Drop the legend .svg in the same folder as the Rmd (or, down the line, the Shiny app)
- Add it to the map using the custom `addLocalLogo()` function below, and avoid `../` to travel backward on the file path.

(Folded code below just loads relevant data + color palette. The legend used in the map is `bivariate_legend.svg` in the `appideas` folder.)
```{r dat, message = F}
library(tidyverse)
library(sf)
library(leaflet)
library(leafem)
library(biscale)
library(patchwork)
library(scales)

places <- read_csv("../data/cdcplaces_cville_tract.csv")
places <- places %>% 
  mutate(GEOID = as.character(locationname))

airquality <- read_csv("../data/airquality_cville_tract.csv")
airquality <- airquality %>% 
  mutate(GEOID = as.character(gid))

tracts <- readRDS("../data/cville_tracts.RDS")

df <- left_join(places, airquality, by = "GEOID")
df <- left_join(tracts, df)

df <- df %>%
  mutate(asthma = ntile(Current_Asthma2018, 3),
         pm25_2016 = ntile(pm2_5_2016, 3)) %>%
  mutate(pm25_2016 = if_else(pm25_2016 == 1, 'A', 
                          if_else(pm25_2016 == 2, 'B', 'C')),
         biclass = paste0(pm25_2016, asthma)) 

bipal <- c("#e8e8e8", "#dfd0d6", "#be64ac", # A-1, A-2, A-3,
           "#ace4e4", "#a5add3", "#8c62aa", # B-1, B-2, B-3
           "#5ac8c8", "#5698b9", "#3b4994") # C-1, C-2, C-2
```

### Leaflet map with legend saved as a local file
Define a new function for calling the locally saved legend:
```{r class.source = 'fold-show'}
# This is a modified version of leafem::addLogo(): https://rdrr.io/cran/leafem/src/R/logo.R
# The HTML that function generated for inserting the locally saved legend had a file path that
# didn't work (at least for me). It also generated and inserted the HTML in two steps by calling
# a separate localImage() function inside of addLogo(). I modified the code to generate HTML with a
# working file path, and, for concision, I condensed things into one function.

addLocalLogo <- function(map,
                    img,
                    alpha = 1,
                    src = c("remote", "local"),
                    url,
                    position = c("topleft", "topright",
                                 "bottomleft", "bottomright"),
                    offset.x = 50,
                    offset.y = 13,
                    width = 60,
                    height = 60) {

  if (inherits(map, "mapview")) map <- mapview2leaflet(map)
  stopifnot(inherits(map, c("leaflet", "leaflet_proxy")))

  position <- position[1]
  src <- src[1]
  
  div_topleft <- paste0("newDiv.css({
                        'position': 'absolute',
                        'top': '", offset.y, "px',
                        'left': '", offset.x, "px',
                        'background-color': 'transparent',
                        'border': '0px solid black',
                        'width': '", width, "px',
                        'height': '", height, "px',
});")
  
  div_topright <- paste0("newDiv.css({
                         'position': 'absolute',
                         'top': '", offset.y, "px',
                         'right': '", offset.x, "px',
                         'background-color': 'transparent',
                         'border': '0px solid black',
                         'width': '", width, "px',
                         'height': '", height, "px',
                         });")
  
  div_bottomleft <- paste0("newDiv.css({
                           'position': 'absolute',
                           'bottom': '", offset.y, "px',
                           'left': '", offset.x, "px',
                           'background-color': 'transparent',
                           'border': '0px solid black',
                           'width': '", width, "px',
                           'height': '", height, "px',
                           });")
  
  div_bottomright <- paste0("newDiv.css({
                            'position': 'absolute',
                            'bottom': '", offset.y, "px',
                            'right': '", offset.x, "px',
                            'background-color': 'transparent',
                            'border': '0px solid black',
                            'width': '", width, "px',
                            'height': '", height, "px',
                            });")
  
  div <- switch(position,
                topleft = div_topleft,
                topright = div_topright,
                bottomleft = div_bottomleft,
                bottomright = div_bottomright)
  
  div_funk <- paste0("function(el, x, data) {
                     // we need a new div element because we have to handle
                     // the mouseover output seperately
                     // debugger;
                     function addElement () {
                     // generate new div Element
                     var newDiv = $(document.createElement('div'));
                     // append at end of leaflet htmlwidget container
                     $(el).append(newDiv);
                     //provide ID and style
                     newDiv.addClass('logo');\n",
                     div,
                     "return newDiv;
                     }")
  
  div_add <- paste0("// check for already existing logo class to not duplicate
                    var logo = $(el).find('.logo');
                    if(!logo.length) {
                    logo = addElement();")

  style <- paste0(', style="opacity:',
                  alpha,
                  ';filter:alpha(opacity=',
                  alpha * 100, ');"')
  
  div_html <- paste0("logo.html('<img src=", '"', img, '"',
                     ", width=", width, ", height=", height, style,
                     ", ></a>');
                       var map = HTMLWidgets.find('#' + el.id).getMap();
                     };
                     }")
  
  render_stuff <- paste0(div_funk, div_add, div_html)
  
  map <- htmlwidgets::onRender(map, render_stuff)
  
  map
}
```

Render the map:
```{r class.source = 'fold-show'}
df_4326 <- st_transform(df, 4326)

factpal <- colorFactor(bipal, domain = df_4326$biclass)

leaflet() %>% 
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = df_4326,
              fillColor = ~factpal(biclass),
              weight = 1,
              opacity = 1,
              color = "white", 
              fillOpacity = 0.8,
              highlight = highlightOptions(
                weight = 2,
                fillOpacity = 0.8,
                bringToFront = T),
              popup = paste0("GEOID: ", df_4326$GEOID, "<br>",
                             "Percent with asthma: ", 
                             df_4326$Current_Asthma2018, "<br>",
                             "PM2.5 rate: ", 
                             round(df_4326$pm2_5_2016,1))) %>%
  addLocalLogo("bivariate_legend.svg", src = "local",
          position = "topleft", width = 100, height = 100,
          alpha = 0.8)
```