---
title: "Script"
author: "Helena Lindsay"
date: "7/7/2021"
output:
  html_document: 
    toc: TRUE
    toc_float: TRUE
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE) 
```


### Data from get_nlcd() 
Data source: https://www.mrlc.gov/

Data was retrieved using the FedData library, get_nlcd function. The function returns a RasterLayer of NLCD data cropped to a given template study area. In this project, we pulled the impervious surfaces and tree canopy data to then analyze based on different tracts, blocks in various locations around Virginia.


```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(leaflet)
library(sp)
library(viridis)
library(raster)
```


## Tree Canopy{.tabset}

### Charlottesville Region {.tabset}

```{r, message=F, warning=F}
cville_tracts <- readRDS("Cville_Tree/cville_tracts.RDS")
cville_blkgps <- readRDS("Cville_Tree/cville_blkgps.RDS")
cville_blocks <- readRDS("Cville_Tree/cville_blocks.RDS")
```

#### Tracts
```{r, message=F, warning=FALSE}
cvl_tracts <- read.csv("Cville_Tree/nlcd_tree_cville_tracts.csv")

head(cvl_tracts)
# cvl_tracts <- cvl_tracts[complete.cases(cvl_tracts),]


cville_tracts$tree_can <- cvl_tracts$tree_can


pal <- colorNumeric("plasma", reverse = TRUE, domain = cvl_tracts$tree_can)
cvl_tracts$tree_can <- as.numeric(cvl_tracts$tree_can)
cville_tracts$COUNTYFP <- as.numeric(cville_tracts$COUNTYFP)
m <- leaflet()%>%
  addTiles()%>%
  addPolygons(data = cville_tracts,
              fillColor = ~pal(tree_can),
              weight = 1,
              opacity = 1,
              color = "white",
              fillOpacity = 0.6,
              highlight = highlightOptions(weight = 2, fillOpacity = 0.8, bringToFront = T),
              popup = paste0("FIPS Code: ", cvl_tracts$GEOID, "<br>",
                             "Tree Canopy: ", cvl_tracts$tree_can)) %>%
  addLegend("bottomright", pal = pal, values = cvl_tracts$tree_can,
            title = "Tree Canopy", opacity = 0.7)
m
```



#### Blockgroups
```{r, message=F, warning=F}
cvl_blkgps <- read.csv("Cville_Tree/nlcd_tree_cville_blkgps.csv")%>%
  as.data.frame()
head(cvl_blkgps)

cville_blkgps$tree_can <- cvl_blkgps$tree_can

pal <- colorNumeric("plasma", reverse = TRUE, domain = cvl_blkgps$tree_can)
cvl_blkgps$tree_can <- as.numeric(cvl_blkgps$tree_can)
cville_blkgps$COUNTYFP <- as.numeric(cville_blkgps$COUNTYFP)
m <- leaflet()%>%
  addTiles()%>%
  addPolygons(data = cville_blkgps,
              fillColor = ~pal(tree_can),
              weight = 1,
              opacity = 1,
              color = "white",
              fillOpacity = 0.6,
              highlight = highlightOptions(weight = 2, fillOpacity = 0.8, bringToFront = T),
              popup = paste0("FIPS Code: ", cvl_blkgps$GEOID, "<br>",
                             "Tree Canopy: ", cvl_blkgps$tree_can)) %>%
  addLegend("bottomright", pal = pal, values = cvl_blkgps$tree_can,
            title = "Tree Canopy", opacity = 0.7)
m

```


#### Blocks
```{r, message=F, warning=FALSE}
cvl_blocks <- read.csv("Cville_Tree/nlcd_tree_cville_blocks.csv")%>%
  as.data.frame()
head(cvl_blocks)

cville_blocks$tree_can <- cvl_blocks$tree_can

pal <- colorNumeric("plasma", reverse = TRUE, domain = cvl_blocks$tree_can)
cvl_blocks$tree_can <- as.numeric(cvl_blocks$tree_can)
cville_blocks$COUNTYFP <- as.numeric(cville_blocks$COUNTYFP)
m <- leaflet()%>%
  addTiles()%>%
  addPolygons(data = cville_blocks,
              fillColor = ~pal(tree_can),
              weight = 1,
              opacity = 1,
              color = "white",
              fillOpacity = 0.6,
              highlight = highlightOptions(weight = 2, fillOpacity = 0.8, bringToFront = T),
              popup = paste0("FIPS Code: ", cvl_blocks$GEOID, "<br>",
                             "Tree Canopy: ", cvl_blocks$tree_can)) %>%
  addLegend("bottomright", pal = pal, values = cvl_blocks$tree_can,
            title = "Tree Canopy", opacity = 0.7)
m
```


### Eastern Shore Region {.tabset}

```{r, message=F, warning=F}
east_tracts <- readRDS("Easternshore_Tree/eastshore_tracts.RDS")
east_blkgps <- readRDS("Easternshore_Tree/eastshore_blkgps.RDS")
east_blocks <- readRDS("Easternshore_Tree/eastshore_blocks.RDS")
```

#### Tracts
```{r, message=F, warning=FALSE}
eastern_tracts <- read.csv("Easternshore_Tree/nlcd_tree_eastshore_tracts.csv")%>%
  as.data.frame()

head(eastern_tracts)
# cvl_tracts <- cvl_tracts[complete.cases(cvl_tracts),]

east_tracts$tree_can <- eastern_tracts$tree_can

pal <- colorNumeric("plasma", reverse = TRUE, domain = eastern_tracts$tree_can)
eastern_tracts$tree_can <- as.numeric(eastern_tracts$tree_can)
east_tracts$COUNTYFP <- as.numeric(east_tracts$COUNTYFP)
m <- leaflet()%>%
  addTiles()%>%
  addPolygons(data = east_tracts,
              fillColor = ~pal(tree_can),
              weight = 1,
              opacity = 1,
              color = "white",
              fillOpacity = 0.6,
              highlight = highlightOptions(weight = 2, fillOpacity = 0.8, bringToFront = T),
              popup = paste0("FIPS Code: ", eastern_tracts$GEOID, "<br>",
                             "Tree Canopy: ", eastern_tracts$tree_can)) %>%
  addLegend("bottomright", pal = pal, values = eastern_tracts$tree_can,
            title = "Tree Canopy", opacity = 0.7)
m
```


#### Blockgroups
```{r, message=F, warning=FALSE}
eastern_blkgps <- read.csv("Easternshore_Tree/nlcd_tree_eastshore_blkgps.csv")%>%
  as.data.frame()

head(eastern_blkgps)
# cvl_tracts <- cvl_tracts[complete.cases(cvl_tracts),]

east_blkgps$tree_can <- eastern_blkgps$tree_can

pal <- colorNumeric("plasma", reverse = TRUE, domain = eastern_blkgps$tree_can)
eastern_blkgps$tree_can <- as.numeric(eastern_blkgps$tree_can)
east_blkgps$COUNTYFP <- as.numeric(east_blkgps$COUNTYFP)
m <- leaflet()%>%
  addTiles()%>%
  addPolygons(data = east_blkgps,
              fillColor = ~pal(tree_can),
              weight = 1,
              opacity = 1,
              color = "white",
              fillOpacity = 0.6,
              highlight = highlightOptions(weight = 2, fillOpacity = 0.8, bringToFront = T),
              popup = paste0("FIPS Code: ", eastern_blkgps$GEOID, "<br>",
                             "Tree Canopy: ", eastern_blkgps$tree_can)) %>%
  addLegend("bottomright", pal = pal, values = eastern_blkgps$tree_can,
            title = "Tree Canopy", opacity = 0.7)
m
```

#### Blocks
```{r, message=F, warning=FALSE}
eastern_blocks <- read.csv("Easternshore_Tree/nlcd_tree_eastshore_blocks.csv")%>%
  as.data.frame()

head(eastern_blocks)
# cvl_tracts <- cvl_tracts[complete.cases(cvl_tracts),]

east_blocks$tree_can <- eastern_blocks$tree_can

pal <- colorNumeric("plasma", reverse = TRUE, domain = eastern_blocks$tree_can)
eastern_blocks$tree_can <- as.numeric(eastern_blocks$tree_can)
east_blocks$COUNTYFP <- as.numeric(east_blocks$COUNTYFP)
m <- leaflet()%>%
  addTiles()%>%
  addPolygons(data = east_blocks,
              fillColor = ~pal(tree_can),
              weight = 1,
              opacity = 1,
              color = "white",
              fillOpacity = 0.6,
              highlight = highlightOptions(weight = 2, fillOpacity = 0.8, bringToFront = T),
              popup = paste0("FIPS Code: ", eastern_blocks$GEOID, "<br>",
                             "Tree Canopy: ", eastern_blocks$tree_can)) %>%
  addLegend("bottomright", pal = pal, values = eastern_blocks$tree_can,
            title = "Tree Canopy", opacity = 0.7)
m
```







