---
title: "Eastern Shore SNAP Data"
author: "Khalila Karefa-Kargbo"
date: "7/27/2021"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(leaflet)
```

## Data Source

Source: ...

* Source URL: ...
* Download URL: ...

```{r}
snap_eastern <- read_csv("~/DemofData/summer-sandbox/eastern_shore_collection/data/snap_eastern.csv")
```

### The Data Used Here
...
explain SNAP data, filtering process, etc

```{r}
knitr::kable(glimpse(snap_eastern), "simple")
```

## Map
```{r}
lon <- pull(snap_eastern, Longitude)
lat <- pull(snap_eastern, Latitude)

PopUpInfo <- paste0(snap_eastern$Store_Name, "<br>",
                    snap_eastern$Address, "<br>",
                    snap_eastern$City, ", ", snap_eastern$State, " ", snap_eastern$Zip5)

leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%  
  setView(lng = mean(lon), lat = mean(lat), zoom = 9) %>% 
  addMarkers(lng = lon, lat = lat, popup = PopUpInfo,
             clusterOptions = markerClusterOptions())
```

## Exploration

use this space to explain some more about the grocery stores, maybe copy and paste script list here

number of grocery stores?

etc
etc
etc

included <- easternstores[c(1, 14:15, 17, 19:20, 33:34, 39:40, 46, 58:59, 65), ] 
14: mexican grocery store
15: el crucero-- fresh produce so including
34: general store but sells produce
65: including bc supermarket in the name, but store doesn't seem to exist? doesn't come up in google search (even when searching address)

