---
title: "Charlottesville SNAP Data"
author: "Khalila Karefa-Kargbo"
date: "7/27/2021"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(leaflet)
```

## Data Source

Source: ...

* Source URL: ...
* Download URL: ...

```{r}
snap_cville <- read_csv("~/DemofData/summer-sandbox/cville_region_collection/data/snap_cville.csv")
meta <- read_sheet("https://docs.google.com/spreadsheets/d/1nqm3DuVXD1ObbVe_deacvT7uSLdBXfQJo3mkbqDwrVo/edit#gid=5733069", sheet = 'snap')
```

### The Data Used Here
...
explain SNAP data, filtering process, etc

```{r}
knitr::kable(glimpse(snap_cville), "simple")
```

## Map
```{r}
lon <- pull(snap_cville, Longitude)
lat <- pull(snap_cville, Latitude)

PopUpInfo <- paste0(snap_cville$Store_Name, "<br>",
                    snap_cville$Address, "<br>",
                    snap_cville$City, ", ", snap_cville$State, " ", snap_cville$Zip5)

leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>%  
  setView(lng = mean(lon), lat = mean(lat), zoom = 9) %>% 
  addMarkers(lng = lon, lat = lat, popup = PopUpInfo,
  clusterOptions = markerClusterOptions())
```

## Exploration

use this space to explain some more about the grocery stores, maybe copy and paste script list here

number of grocery stores
etc
etc
etc

included <- stores_4326[c(4, 6, 10:13, 19:20, 33:34, 44:45, 47:49, 52:53, 56:58,
                          65, 74, 76, 78, 80, 84, 88:89, 92, 94, 101, 105:107, 
                          111:112, 114:115, 117, 125, 127, 144, 146, 156, 161:163,
                          165, 167, 169), ] 
20: international food grocery stores so may not fit criteria of "wide variety" of healthy foods, but does for some people so going to include
44: african market place ^
49: Rebecca's Natural Food-- definitely not a large grocery store but it does provide healthy food options?
57: integral yoga-- same issue ^
65: asian market-- int'l food...
112: La Guadalupana ^
127: indian bazar-- int'l food
161: latin market...
169: indian & nepali

