---
title: "National Flood Hazard Layer Data"
author: "Chase Dawson, Michele Claibourn"
date: "11/30/2021"
output:
  html_document: 
    toc: TRUE
    toc_float: TRUE
    code_folding: hide
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE) 
```


```{r}
# Load libraries
library(tidyverse)
library(leaflet)
library(sf)
library(viridis)
library(raster)
library(stargazer)

# Load data
eastern_tracts <- readRDS("../data/eastshore_tracts.RDS")
eastern_blkgps <- readRDS("../data/eastshore_blkgps.RDS")
eastern_blocks <- readRDS("../data/eastshore_blocks.RDS")

nfhl_eastern_tracts <- read_csv("../dataraw/nfhl_eastern_tracts.csv")
nfhl_eastern_blkgps <- read_csv("../dataraw/nfhl_eastern_blkgps.csv")
nfhl_eastern_blocks <- read_csv("../dataraw/nfhl_eastern_blocks.csv")

# arrange data
nfhl_eastern_tracts <- nfhl_eastern_tracts %>% 
  dplyr::select(-c(`...1`, spatial_unit)) %>% 
  mutate(GEOID = as.character(GEOID))

nfhl_eastern_tracts2 <- nfhl_eastern_tracts %>% 
  pivot_wider(names_from = zone, values_from = c(area, perc))

nfhl_eastern_blkgps <- nfhl_eastern_blkgps %>% 
  dplyr::select(-c(`...1`, spatial_unit)) %>% 
  mutate(GEOID = as.character(GEOID))

nfhl_eastern_blkgps2 <- nfhl_eastern_blkgps %>% 
  pivot_wider(names_from = zone, values_from = c(area, perc))

nfhl_eastern_blocks <- nfhl_eastern_blocks %>% 
  dplyr::select(-c(`...1`, spatial_unit)) %>% 
  mutate(GEOID = as.character(GEOID))

nfhl_eastern_blocks2 <- nfhl_eastern_blocks %>% 
  pivot_wider(names_from = zone, values_from = c(area, perc))

# Join data to geometry
nfhl_eastern_tracts_geo <- left_join(eastern_tracts, nfhl_eastern_tracts2)
nfhl_eastern_blkgps_geo <- left_join(eastern_blkgps, nfhl_eastern_blkgps2)
nfhl_eastern_blocks_geo <- left_join(eastern_blocks, nfhl_eastern_blocks2, by = c("GEOID10" = "GEOID"))
```

## Data Source

* The data represented here come from FEMA's [National Flood Hazard Layer](https://www.fema.gov/flood-maps/national-flood-hazard-layer), containing current effective flood hazard data, as downloaed through [FEMA Flood Map Service Center.](https://msc.fema.gov/portal/advanceSearch#searchresultsanchor) 
* The current data was downloaded in November 2021; the NFHL is being revised continuously.


## Variable Descriptions and Summaries

Variables include: 

```{r}
glimpse(nfhl_eastern_tracts2)
```

* GEOID: geographic identifer for spatial area (tracts, block groups, and blocks)
* area_0: area (in km^2) of block/block group/tract/county that do not intersect with 100-year or 200-year floodplains, or areas with minimal chance of flooding.
* area_1per: area (in km^2) of block/block group/tract/county that intersects with 100-year floodplain, or areas with a 1% annual chance of flooding.
* area_2per: area (in km^2) of block/block group/tract/county that intersects with 100-year floodplain, or areas with a 2% annual chance of flooding. This zone is also used to designate base floodplains of lesser hazards, such as areas protected by levees from 100-year flood, or shallow flooding areas with average depths of less than one foot or drainage areas less than 1 square mile.
* perc_0: percentage of block/block group/tract/county that does not overlap with the 1% or 2% annual flood zone.
* perc_1per: percentage of block/block group/tract/county that overlaps with the 1% annual flood zone.
* perc_2per: percentage of block/block group/tract/county that overlaps with the 2% annual flood zone.

5-number summaries of variables by block groups:

```{r}
nfhl_eastern_blkgps2 %>%  
  dplyr::select(where(~is.numeric(.x) && !is.na(.x))) %>% 
  as.data.frame() %>% 
  stargazer(., type = "text", title = "Summary Statistics", digits = 1,
            summary.stat = c("mean", "sd", "min", "median", "max"))
```

## Visual Distributions{.tabset}

### Tracts
Percent of tract in minimal flood, 1% annual floodplain, and 2% annual floodplain zones
```{r}
nfhl_eastern_tracts %>% 
  ggplot(aes(x = perc, fill = zone)) + 
  geom_histogram() + 
  scale_fill_viridis(option = "viridis", discrete = TRUE, guide = FALSE) +
  facet_wrap(~zone, scales = "free")
```
  
### Block Groups
Percent of block group in minimal flood, 1% annual floodplain, and 2% annual floodplain zones
```{r}
nfhl_eastern_blkgps %>% 
  ggplot(aes(x = perc, fill = zone)) + 
  geom_histogram() + 
  scale_fill_viridis(option = "viridis", discrete = TRUE, guide = FALSE) +
  facet_wrap(~zone, scales = "free")
```

### Blocks 
Percent of block in minimal flood, 1% annual floodplain, and 2% annual floodplain zones
```{r}
nfhl_eastern_blocks %>% 
  ggplot(aes(x = perc, fill = zone)) + 
  geom_histogram() + 
  scale_fill_viridis(option = "viridis", discrete = TRUE, guide = FALSE) +
  facet_wrap(~zone, scales = "free")
```

## Maps
### 1% Annual Flood Zone{.tabset}

#### Tracts
```{r}
pal <- colorNumeric("viridis", reverse = TRUE, domain = nfhl_eastern_tracts_geo$perc_1per)

leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = nfhl_eastern_tracts_geo,
              fillColor = ~pal(perc_1per),
              weight = 1,
              opacity = 1,
              color = "white",
              fillOpacity = 0.6,
              highlight = highlightOptions(weight = 2, fillOpacity = 0.8, bringToFront = T),
              popup = paste0("FIPS Code: ", nfhl_eastern_tracts_geo$GEOID, "<br>",
                             "Percent in 1% floodplain: ", nfhl_eastern_tracts_geo$perc_1per)) %>%
  addLegend("bottomright", pal = pal, values = nfhl_eastern_tracts_geo$perc_1per,
            title = "Percent in 1% Annual Floodplain", opacity = 0.7)
```


#### Block Groups
```{r}
pal <- colorNumeric("viridis", reverse = TRUE, domain = nfhl_eastern_blkgps_geo$perc_1per)

leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = nfhl_eastern_blkgps_geo,
              fillColor = ~pal(perc_1per),
              weight = 1,
              opacity = 1,
              color = "white",
              fillOpacity = 0.6,
              highlight = highlightOptions(weight = 2, fillOpacity = 0.8, bringToFront = T),
              popup = paste0("FIPS Code: ", nfhl_eastern_blkgps_geo$GEOID, "<br>",
                             "Percent in 1% floodplain: ", nfhl_eastern_blkgps_geo$perc_1per)) %>%
  addLegend("bottomright", pal = pal, values = nfhl_eastern_blkgps_geo$perc_1per,
            title = "Percent in 1% Annual Floodplain", opacity = 0.7)
```

#### Blocks

Block level maps are slow to render and may be added later.


### 2% Annual Flood Zone{.tabset}

#### Tracts
```{r}
pal <- colorNumeric("viridis", reverse = TRUE, domain = nfhl_eastern_tracts_geo$perc_2per)

leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = nfhl_eastern_tracts_geo,
              fillColor = ~pal(perc_2per),
              weight = 1,
              opacity = 1,
              color = "white",
              fillOpacity = 0.6,
              highlight = highlightOptions(weight = 2, fillOpacity = 0.8, bringToFront = T),
              popup = paste0("FIPS Code: ", nfhl_eastern_tracts_geo$GEOID, "<br>",
                             "Percent in 2% floodplain: ", nfhl_eastern_tracts_geo$perc_2per)) %>%
  addLegend("bottomright", pal = pal, values = nfhl_eastern_tracts_geo$perc_2per,
            title = "Percent in 2% Annual Floodplain", opacity = 0.7)
```

#### Block Groups
```{r}
pal <- colorNumeric("viridis", reverse = TRUE, domain = nfhl_eastern_blkgps_geo$perc_2per)

leaflet() %>%
  addProviderTiles("CartoDB.Positron") %>% 
  addPolygons(data = nfhl_eastern_blkgps_geo,
              fillColor = ~pal(perc_2per),
              weight = 1,
              opacity = 1,
              color = "white",
              fillOpacity = 0.6,
              highlight = highlightOptions(weight = 2, fillOpacity = 0.8, bringToFront = T),
              popup = paste0("FIPS Code: ", nfhl_eastern_blkgps_geo$GEOID, "<br>",
                             "Percent in 2% floodplain: ", nfhl_eastern_blkgps_geo$perc_2per)) %>%
  addLegend("bottomright", pal = pal, values = nfhl_eastern_blkgps_geo$perc_2per,
            title = "Percent in 2% Annual Floodplain", opacity = 0.7)
```

#### Blocks

Block level maps are slow to render and may be added later.
