# Individual Household Program

FAQ Sheet:
<file:///Users/khalila/Downloads/Individual%20and%20Households%20Program%20Remote%20Inspections.pdf>

```{r}
library(tidyverse)
```

"IHP is intended to meet basic needs and supplement disaster recovery
efforts. The IHP is not intended to return disaster-damaged property to
its pre-disaster condition. Disaster damage to secondary or vacation
homes does not qualify for IHP assistance"

"FEMA's Individuals and Households Program (IHP) provides financial and
direct services to eligible individuals and households affected by a
disaster, who have uninsured or under-insured necessary expenses and
serious needs. IHP assistance is not a substitute for insurance and
cannot compensate for all losses caused by a disaster. The assistance is
intended to meet your basic needs and supplement disaster recovery
efforts."

"Any financial information is derived from NEMIS and not FEMA's official
financial systems. Due to differences in reporting periods, status of
obligations and how business rules are applied, this financial
information may differ slightly from official publication on public
websites such as usaspending.gov; this dataset is not intended to be
used for any official federal financial reporting."

**Get Data**

```{r}
ihp <- read.csv("~/DemofData/IndividualsAndHouseholdsProgramValidRegistrations.csv", header=TRUE)
```

# Notes

only spatial units available are **county, city, state** ie. no
tract/FIPS

**Potentially relevant variables:**

-   incidentType: Type of incident such as fire or flood. The incident type will affect the types of assistance available.
    - after looking at data, the only incident type in our regions is Hurricanes

-   declarationDate: Date the disaster was declared
    - declaration date for every single observation in both relevant
        regions is 2003-09-19 : must be referring to Hurricane Isabel

-   grossIncome:Self-reported gross income range in U.S. dollars

-   residenceType: Damaged Dwelling Residence Type (e.g., single-family home, apartment) 

-   floodInsurance: Does the qpplicant have flood insurance? 
  -    44/469 (9.38%) had insurance in eastern shore 
  -    11/508 (2.17%) in cville 

-   fipAmount: The amount of the applicant's Flood Insurance Premium (FIP) that was paid for by FEMA, in U.S. dollars

-   haAmount: Total amount awarded for Housing Assistance under IHP

-   onaAmount: Total dollar amount awarded for Other Needs Assistance (ONA) from the Individual and Households Program (IHP) 

-   floodDamage: Was flood damage recorded?

-   floodDamageAmount: FEMA-determined amount of damage to real and personal property due to flood damage, in U.S. dollars

-   rentalAssistanceAmount: Amount of Rental Assistance awarded in U.S. dollars 
  -   eligibility variable: rentalAssistanceEligible 
  
-   repairAmount: Amount of Repair Assistance awarded in U.S. dollars

-   personalPropertyAmount: Amount of ONAnawarded for Personal Property Assistance in U.S. dollars
  -   eligibility variable: personalPropertyEligible

According to FEMA, the incident types are: Biological, Chemical,
**Coastal Storm**, Dam/Levee Break, **Drought**, Earthquake, Explosion, Fire, Fishing Losses, **Flood**, Freezing, Human Cause, **Hurricane**, Mud/Landslide, Other, Severe Ice Storm, **Severe Storm(s)**, Snow, Terrorist, Tornado, Toxic Substances, Tsunami, Typhoon, Volcano

```{r}
eastern_ihp <- ihp %>% 
  filter(damagedStateAbbreviation == 'VA',
         county %in% (c('Accomack (County)', 'Northampton (County)')))
# 469 observations

cville_ihp <- ihp %>%
  filter(damagedStateAbbreviation == 'VA',
         county %in% (c('Charlottesville', 'Albemarle (County)','Fluvanna (County)', 'Greene (County)', 'Louisa (County)', 'Nelson (County)')))
# 508 observations
```

```{r}
ihpVA <- ihp %>%
  filter(damagedStateAbbreviation == 'VA')
```

The only incident type in both regions is **'Hurricanes'**. This does
seem a bit odd to me though, there's never been any severe storms or
coastal floodings on the Eastern Shore?
