# Individuals and Households Program

Source: FEMA, Individuals and Households Program, last updated June 2021
* Source URL: [https://www.fema.gov/openfema-data-page/individuals-and-households-program-valid-registrations-v1]

FAQ Sheet:
<file:///Users/khalila/Downloads/Individual%20and%20Households%20Program%20Remote%20Inspections.pdf>

```{r}
library(tidyverse)
library(stargazer)
```

"IHP is intended to meet basic needs and supplement disaster recovery
efforts. The IHP is not intended to return disaster-damaged property to
its pre-disaster condition. Disaster damage to secondary or vacation
homes does not qualify for IHP assistance"

"FEMA's Individuals and Households Program (IHP) provides financial and
direct services to eligible individuals and households affected by a
disaster, who have uninsured or under-insured necessary expenses and
serious needs. IHP assistance is not a substitute for insurance and
cannot compensate for all losses caused by a disaster. The assistance is
intended to meet your basic needs and supplement disaster recovery
efforts."

"Any financial information is derived from NEMIS and not FEMA's official
financial systems. Due to differences in reporting periods, status of
obligations and how business rules are applied, this financial
information may differ slightly from official publication on public
websites such as usaspending.gov; this dataset is not intended to be
used for any official federal financial reporting."

**Get Data**

```{r}
ihp <- read.csv("~/DemofData/IndividualsAndHouseholdsProgramValidRegistrations.csv", header=TRUE)
```

```{r}
ihpVA <- ihp %>%
  filter(damagedStateAbbreviation == 'VA')
```

```{r}
eastern_ihp <- ihp %>% 
  filter(damagedStateAbbreviation == 'VA',
         county %in% (c('Accomack (County)', 'Northampton (County)')))
# 469 observations

cville_ihp <- ihp %>%
  filter(damagedStateAbbreviation == 'VA',
         county %in% (c('Charlottesville', 'Albemarle (County)','Fluvanna (County)', 'Greene (County)', 'Louisa (County)', 'Nelson (County)')))
# 508 observations
```

```{r}
glimpse(ihp)
```

# Notes

only spatial units available are **county, city, state** ie. no tract/FIPS

### Potentially relevant variables:
(fairly long list; I think they key ones are rpfvl and ppfvl)

-   incidentType: Type of incident such as fire or flood. The incident type will affect the types of assistance available.
    - after looking at data, the only incident type in our regions is Hurricanes

-   declarationDate: Date the disaster was declared
    - declaration date for every single observation in both relevant regions is 2003-09-19 : must be referring to Hurricane Isabel

-   floodInsurance: Does the qpplicant have flood insurance? 
  -  44/469 (9.38%) had insurance in eastern shore 
  -  11/508 (2.17%) in cville 

-   fipAmount: The amount of the applicant's Flood Insurance Premium (FIP) that was paid for by FEMA, in U.S. dollars
```{r}
stargazer(data.frame('fipAmount' = cville_ihp[, c('fipAmount')]),
          summary = T, type = 'text', flip = T,
          summary.stat = c('mean', 'sd', 'min', 'max'))

stargazer(data.frame('fipAmount' = eastern_ihp[, c('fipAmount')]),
          summary = T, type = 'text', flip = T,
          summary.stat = c('mean', 'sd', 'min', 'max'))
```

-   haAmount: Total amount awarded for Housing Assistance under IHP
```{r}
stargazer(data.frame('haAmount' = cville_ihp[, c('haAmount')]),
          summary = T, type = 'text', flip = T,
          summary.stat = c('mean', 'sd', 'min', 'max'))

stargazer(data.frame('haAmount' = eastern_ihp[, c('haAmount')]),
          summary = T, type = 'text', flip = T,
          summary.stat = c('mean', 'sd', 'min', 'max'))
```

-   onaAmount: Total dollar amount awarded for Other Needs Assistance (ONA) from the Individual and Households Program (IHP) 
  - eligibility variable: onaEligible
```{r}
stargazer(data.frame('onaAmount' = cville_ihp[, c('onaAmount')]),
          summary = T, type = 'text', flip = T,
          summary.stat = c('mean', 'sd', 'min', 'max'))

stargazer(data.frame('onaAmount' = eastern_ihp[, c('onaAmount')]),
          summary = T, type = 'text', flip = T,
          summary.stat = c('mean', 'sd', 'min', 'max'))
```
  
-   floodDamage: Was flood damage recorded?
  - present in the data dictionary online but not in the csv?
  
-   floodDamageAmount: FEMA-determined amount of damage to real and personal property due to flood damage, in U.S. dollars
  - same as above, it's online in the data dictionary but doesn't seem to be in the csv :/
  
-   rpfvl: FEMA-determined value of disaster-caused damage to real property components, including floors, walls, access roads and bridges, electrical, plumbing, HVAC, etc. Note: IHP does not address land damage.	
```{r}
stargazer(data.frame('rpfvl' = cville_ihp[, c('rpfvl')]),
          summary = T, type = 'text', flip = T,
          summary.stat = c('mean', 'sd', 'min', 'max'))

stargazer(data.frame('rpfvl' = eastern_ihp[, c('rpfvl')]),
          summary = T, type = 'text', flip = T,
          summary.stat = c('mean', 'sd', 'min', 'max'))
```
-   ppfvl: FEMA-determined value of disaster-caused damage to personal property components, including appliances, furniture, etc. Note: IHP does not address land damage.	
```{r results='asis'}
stargazer(data.frame('ppfvl' = cville_ihp[, c('ppfvl')]),
          summary = T, type = 'text', flip = T,
          summary.stat = c('mean', 'sd', 'min', 'max'))

stargazer(data.frame('ppfvl' = eastern_ihp[, c('ppfvl')]),
          summary = T, type = 'text', flip = T,
          summary.stat = c('mean', 'sd', 'min', 'max'))
```

-   rentalAssistanceAmount: Amount of Rental Assistance awarded in U.S. dollars 
  - eligibility variable: rentalAssistanceEligible 
```{r results='asis'}
stargazer(data.frame('rentalAssistanceAmount' = cville_ihp[, c('rentalAssistanceAmount')]),
          summary = T, type = 'text', flip = T,
          summary.stat = c('mean', 'sd', 'min', 'max'))

stargazer(data.frame('rentalAssistanceAmount' = eastern_ihp[, c('rentalAssistanceAmount')]),
          summary = T, type = 'text', flip = T,
          summary.stat = c('mean', 'sd', 'min', 'max'))
```
-   repairAmount: Amount of Repair Assistance awarded in U.S. dollars
```{r}
stargazer(data.frame('repairAmount' = cville_ihp[, c('repairAmount')]),
          summary = T, type = 'text', flip = T,
          summary.stat = c('mean', 'sd', 'min', 'max'))

stargazer(data.frame('repairAmount' = eastern_ihp[, c('repairAmount')]),
          summary = T, type = 'text', flip = T,
          summary.stat = c('mean', 'sd', 'min', 'max'))
```

-   personalPropertyAmount: Amount of ONA awarded for Personal Property Assistance in U.S. dollars
  - eligibility variable: personalPropertyEligible
```{r results='asis'}
stargazer(data.frame('personalPropertyAmount' = cville_ihp[, c('personalPropertyAmount')]),
          summary = T, type = 'text', flip = T,
          summary.stat = c('mean', 'sd', 'min', 'max'))

stargazer(data.frame('personalPropertyAmount' = eastern_ihp[, c('personalPropertyAmount')]),
          summary = T, type = 'text', flip = T,
          summary.stat = c('mean', 'sd', 'min', 'max'))
```
***

According to FEMA, the incident types are: Biological, Chemical, **Coastal Storm**, Dam/Levee Break, **Drought**, Earthquake, Explosion, Fire, Fishing Losses, **Flood**, Freezing, Human Cause, **Hurricane**, Mud/Landslide, Other, Severe Ice Storm, **Severe Storm(s)**, Snow, Terrorist, Tornado, Toxic Substances, Tsunami, Typhoon, Volcano


The only incident type in both regions is **'Hurricanes'**. This doe seem a bit odd to me though, there's never been any severe storms or coastal floodings in the Eastern Shore?
